### 2014-08-05 frontback/jury.tex

% Jury ====================================================================

\pdfbookmark[1]{Jury members}{Jury members}
\chapter*{Jury members}

\noindent \textsc{David Posada}, Professor at the University of Vigo, Vigo, Spain; \\

\noindent Rest of them swell fellas; \\