### 2014-09-14 frontback/summary.tex
improving the General Intro

% Summary ====================================================================

\pdfbookmark[1]{Summary}{Summary}
\chapter*{Summary}

TODO

28 grams of prevention are worth 454 grams of cure
-- Old saying