### 2014-10-19 chapters/chapter01.tex
corrections

\chapter{Introduction}

\begin{remark}{Outline}
In this chapter, we introduce fundamental concepts of the molecular evolution and describe how they are integrated into the field of phylodynamics, a discipline that describes the interaction of evolutionary and ecological dynamics underlying pathogen spread.
Different viral pathogens that emerged in recent history illustrate the vulnerability of the human population to pandemic threats.
We stress the importance of developing efficient and statistically sound surveillance methods that can assist in characterizing those threats.

We subsequently describe some of the most important breakthroughs in the history of molecular evolution, i.e. the introduction of a molecular clock, which allows dating divergence events from genetic data and historical calibration points or sampling times, the shift of the field towards probabilistic methods, the advent of Monte Carlo algorithms as well as powerful and increasingly more attainable computer hardware, which facilitated the adoption of Bayesian approaches.
We further explain how these advances - implemented into publicly available software - adequately account for phylogenetic uncertainty by averaging over all plausible evolutionary histories.

Finally Subsection~\ref{sub:objectives} positions the work presented in this thesis and outlines how it advances our understanding of the processes shaping viral evolution and spread.
The additional sections of this introductory chapter are devoted towards describing the most basic inference methodology in more details, and constitute the background knowledge for the work presented in Chapters~\ref{chap:epoch}--\ref{chap:pibuss}.
\end{remark}

\section{General Introduction}

\subsection{Viral Epidemiology and the Emergence of Phylodynamics\label{sub:epidemiology}}

Many pathogens represent a clear and eminent threat to public health. 
This has been exemplified by 
the outbreak of the Severe Acute Respiratory Syndrome (SARS) in China \citep{Ksiazek2003},
the H1N1 swine-origin influenza A virus outbreak in 2009 \citep{Fraser2009},
which attained pandemic status in only a few months time, but also more recently by a large outbreak of the Ebola haemorrhagic fever in humans in West Africa \citep{Dudas2014}.

Many studies on emerging infectious diseases (EIDs) have highlighted viral pathogens as a major threat to the public health, owing to their high rates of nucleotide substitution, and therefore high capacity to adapt to new hosts. 
Due to highly intensified human mobility there are virtually no natural barriers for the spread of human RNA viruses \citep{Brockmann2006}.
In the recent decades EIDs are becoming increasingly widespread, with their impact and spread only rising over time \citep{Jones2008}.
This may not be surprising since it is well-acknowledged that disease emergence is largely a product of anthropogenic and demographic changes, making it a hidden cost of human economic development.

Combating viral spread and its associated disease burden is a tremendous challenge requiring significant research efforts and decided public health measures. 
Viral sequence data has been recognized as a major asset in the characterization of pathogens that threaten public health.
Molecular epidemiology complements traditional epidemiological studies because it primarily focuses on the etiological agent rather than the host \citep{Leitner2002}.
In particular, the historical information contained in viral gene sequences contributes to a better insight in emergence and early transmission dynamics, even before systematic epidemiological surveillance has been initiated. 
For many years, phylogenetic analysis has been the primary tool in molecular epidemiology. 
By reconstructing a phylogenetic tree, we are able to study the evolutionary history of viral strains, assess epidemiological linkage amongst them, and shed light on transmission patterns. 
With the advent of high-throughput DNA sequencing, genomic data from RNA viruses is becoming available in unprecedented quantities and with remarkable rapidity.

The last decade has witnessed a particular focus of analyses of viral evolutionary dynamics on integrating concepts of phylogenetic and epidemiological dynamics, a research direction originally coined `phylodynamics' \citep{Grenfell2004}.
By aiming at the development of a full quantitative understanding of the processes that shape the epidemiology and evolution of RNA viruses, %PL: please add the Grenfell and Holmes 2009 Plos Com Biol reference for this statement
phylodynamics has become a burgeoning area of research.
In order to perform comprehensive and quantitative analyses of the interaction between the ecological and evolutionary dynamics of RNA viruses, new computational developments are needed and some of these are actively bridging the gap with the field of mathematical epidemiology.

Advances in epidemiological modeling have made it possible to describe the transmission of contagions through individuals.
Perhaps the simplest and best-studied model is the ${\mathcal{S} \rightarrow \mathcal{I} \rightarrow \mathcal{R}}$  compartmental model % TODO:c itation
, which describes how the (S)usceptible individuals become (I)nfected and the (R)ecover, as a function of time.  
These models can be extended to include birth and death events % TODO: citation
and are instrumental in finding the number of secondary infections caused by a single infected individual in an entirely susceptible population, a basic reproductive quantity denoted by $R_{0}$.
This quantity determines whether and epidemic occurs or whether the diseases will die out.
Traditionally, parameters in these models were inferred using surveillance data, %yet more complicated epidemiological studies involving e.g. seasonality, multiple populations with varying $R_0$ or other possible epidemiological predictors are often analytically intractable.
%In that light, developments to infer parameters of compartmental models by studying 
but increasingly pathogen genetic data is being used to inform SIR model parameters \citep{Volz2009}.
%molecular sequence data have become increasingly important \citep{Volz2009}.
Also patterns of geographic diffusion of infectious diseases can be inferred from sequence data \citep{Lemey2009}, sometimes in combination with information on the host movement \citep{Lemey2014}. %PL: add our 2014 GLM-diffusion reference
Integrating population structure into compartmental models which explicitly track the transmission between the infected individuals can provide new avenues for inferring patterns of viral spread. %PL: I think there is a very recent paper by Rasmussen, Volz and Koelle on this

Whereas for emerging pathogens modeling efforts are typically focused on providing estimates of the basic reproduction number, for ongoing epidemics one is typically interested in quantifying the effects of control and preventive measures on between-host scales and the effect of treatment on within-host scale.
In that respect it is important to characterize adaptive evolution, 
%For that reason most efforts are focused on reliable detection of adaptive selection pressure,
a major driving force of viral evolution.
\cite{Volz2013} report nine Influenza A H1N1 vaccine updates being recommended by the World Health Organization (WHO) between 1978 and 2009, as compared to 20 updates recommended for Influenza H3N2, suggesting that the latter undergoes much more adaptive evolution that the former.
Unfortunately positive selection is notoriously difficult to detect, as on the sequence level it may apply to only a handful of sites, on the tree topology level it may apply to only some of the branches, or even only to a particular time on that branch.
A substantial amount of work has been aimed at developing models which can detect selection acting differently between sites, between branches or between a combination of those \citep{NY98, Pond2005a, Goode2008}.

In order to position the work presented here, we first provide a brief general account of the field of molecular evolution and phylogenetics and then introduce recent developments and models specific for rapidly-evolving RNA viruses, and how this culminated in to a state-of-the-art statistical framework for phylodynamics.
As a final part of the general introduction, we outline how the current doctoral thesis work extends and complements this framework in the objectives.
The remainder of the chapter provides a more mathematically-oriented description of how to model evolution evolution as a stochastic process on trees, the computational aspects involved, and how to treat the problem from a Bayesian statistical perspective.

\subsection{Molecular evolution and Phylogenetics\label{sub:molevol}}

%PL: got confused an rewrote this.
%The main axiom of molecular evolution is the idea that the processes that shape the fitness of organisms leave a measurable imprint on their genomes.
%Evolution is a stochastic process that can introduce at any given moment random base changes in the genetic alphabet.
%For some time those mutations coexist with the vanilla-type %PL: what do you mean with vanilla-type?
%organisms until selection decides whether the mutations increases the fitness, in which case the entire population eventually carries the said mutation and it becomes a fixed substitution.

Evolution is the product of mutational processes that stochastically introduce changes into the genetic make of organisms and population processes that lead to the elimination, fixation or partial maintenance of these mutations in the population.
The latter occurs through genetic drift or the more deterministic process of natural selection. 
Adaptation by natural selection is the most important process in biology as it determines the fate of mutations that impact the fitness of the organisms within their variable environment.

Molecular phylogenetics studies the evolutionary pathways underlying these processes that lead in the polymorphism data observed in contemporary times.
Based on the sequence data obtained for a number of organisms or taxa, molecular phylogenetics aims at inferring the ancestral relationships and evolutionary parameters of interest.
The result of a molecular phylogenetic analysis is represented in a phylogenetic tree. 
Phylogenetic trees are bifurcating graphs depicting the pattern of common ancestry among the taxa.
External nodes, tips or leafs in the tree represent the observed character data, while internal nodes represent the missing data for hypothetical ancestors.  
The common ancestor of all the observed character sequences (the most-recent common ancestor, MRCA) is referred to the as the root of the tree.
In unrooted trees, the position of the root is unknown, and unlike rooted trees, they are therefore lacking an evolutionary direction.
A group of taxa in a phylogeny, consisting of all the descendants sharing a particular ancestor, is called a clade or phylogenetic cluster.

%---FIRST TREES---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{first_trees} 
\caption{
{ \footnotesize 
{\bf Phylogenetic trees.} A. A page from Darwin's notebooks around July 1837 showing his first sketch of an evolutionary tree.
B. First phylogeny estimated from genetic data and published by \cite{Fitch1967}.
}% END: footnotesize
}
\label{fig:first_trees}
\end{figure}

%Underlying assumption of the phylogenetic tree is that the data used to infer them indeed encompasses some sort of evolutionary relatedness.
Phylogenetic thinking traces back to the first know figure of a phylogenetic tree, sketched by Charles Darwin in %1937 
1837
on the margin side of his notebook (see Figure \ref{fig:first_trees}), with the famous words \emph{`I think'} written above it.
In those days there was no notion of genetic data, therefore Darwin conceived his trees by comparing morphology, assuming that organisms with similar forms and shapes descended from more recent common ancestors. %\footnote{Furry with furry.}.
Modern phylogenetics use molecular data and a combination of statistical and computational methods to reconstruct phylogenetic trees.
%distance-based trees
Among the first tractable methods to reconstruct trees from molecular data were the hierarchical clustering methods based on genetic distance \citep{Fitch1967}, defined as the number of nucleotides that would need to be altered to make two sequences identical (see Figure \ref{fig:first_trees} B).
These methods were criticized for not making explicit use of the relationship between individual characters of the genetic data, but rather reducing it to distance based measures.
%Moreover by using pairwise comparisons some of the information is inevitably lost.
Moreover, hierarchical clustering leads to a single point estimate of the evolutionary history and therefore does not consider alternative hypotheses for the data.

% parsimony
In order to evaluate different tree topologies the idea or parsimony was adopted in phylogenetics. 
Maximum parsimony approaches attempt to find the tree that requires minimum number of evolutionary changes to explain the data. %, maing use of the whole data at once.
This concept stems from the work of William of Ockham (1320), who stated that among competing hypotheses one should select the simplest one and disregard the remaining more complex ones (the Ockham's razor).
%PL: It was really Cavalli-Sforza and Edwards, both students from Fisher, who introduced the idea into phylogenetics. Indeed, Fitch then introduced a tractable algorithm to compute parsimony scores. They also introduced likelihood and in the same way, Felsenstein, then introduced the tractable pruning algorithm to compute likelyhoods. It would be good to find the reference of their paper and add it here as well below in the first sentence on likelihood as both ideas were mentioned in the same paper.
\cite{Fitch1971} published an dynamic programming algorithm that efficiently searches all possible tree topologies, assigning scores to them and finding the one with minimal parsimony score.
Figure \ref{fig:parsimony} presents an example of parsimony, where among two trees explaining the evolution of sequences {\color{green}AAG}, {\color{green}AAA}, {\color{green}GGA}, {\color{green}AGA} only the most parsimonious one is selected.

%---PARSIMONY TREE---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.3]{parsimony} 
\caption{
{ \footnotesize 
{\bf Parsimony example.} The tree on the left is preferred as it requires the least amount changes (3 substitutions as opposed to 4 for the tree on the right).
}% END: footnotesize
}
\label{fig:parsimony}
\end{figure}

%likelihood
As a statistical alternative to an optimality criterion, likelihood had also been proposed for phylogenetic inference relatively early.
Maximum likelihood is a probabilistic method of inference, that evaluates a hypothesis in terms of how probable it is in the light of the observed data.
In phylogenetic setting, a likelihood function returns the probability that the proposed model and the proposed tree, understood as a topology and branch lengths, have generated the observed data.
The evolutionary history and associated set of evolutionary parameters with a higher probability is naturally preferred over a history with a lower probability.
By using optimization techniques one can find the parameter values that maximize the probability of observing the data.
In his influential work, \cite{Felsenstein1981} proposed an efficient algorithm for calculating a likelihood of sequence evolution on a tree under an evolutionary model of substitution, the so called `tree pruning' algorithm.
Unfortunately finding the maximum likelihood tree requires a search through the space of all possible tree topologies; in section \ref{sub:mlburden} we discuss in detail the computational burden associated with these calculations.
The time required to exhaustively search through a space of all possible rooted trees grows explosively with the number of taxa, as depicted in Table \ref{tab:maxLikeBurden}.

\begin{table}[H]
\centering
% \footnotesize{
\begin{tabular}{cc}
% {@{}l*{2}{D{.}{.}{7}}@{}}
\hline 
Tips & Rooted trees $\frac{(2N-3)!}{2^{N-2}\cdot(N-2)!}$ \tabularnewline
\hline 
\rowcolor{gray1}
$3$ & $3$\tabularnewline
$6$ & $945$\tabularnewline
\rowcolor{gray1}
$9$ & $2'027'025$\tabularnewline
$15$ & $2\times10^{14}$\tabularnewline
\rowcolor{gray1}
$20$ & $8\times10^{21}$\tabularnewline
$55$ & $3.19\times10^{86}$\tabularnewline
\end{tabular}
% }
\caption{
{ \footnotesize 
{\bf{Number of rooted trees per number of taxa.}} For comparison, $4 \times 10^{80} $ is the number of atoms in the observable universe.
}% END: footnotesize
}
\label{tab:maxLikeBurden}
\end{table}

%bayesian
The advent of powerful and increasingly more attainable compute power has been instrumental to the adoption of statistical methods based on likelihood estimation and the closely related Bayesian methods.
Bayesian statistical thinking is  not a new idea and stems back to the 18th century work of reverend Thomas Bayes, a statistician, philosopher and theologian.

%---BAYES---%
\begin{remark}{Thomas Bayes}
\begin{figure}[H]
\centering
\includegraphics[scale=0.4]{Thomas_Bayes} 
\caption{
{ \footnotesize 
{\bf Only known portrait of reverend Thomas Bayes.} This portrait was published in the 1936 book on the history of actuarial science `History of Life Insurance'. It is only hypothesised to be an actual portrait of the scholar.
}% END: footnotesize
}
\label{fig:bayes}
\end{figure}
\end{remark}

Interestingly Bayes has never published his famous work.
Bayesian statistics have always been appealing as a very flexible approach to inference, as they allow for inclusion of prior belief about any unknown parameters, which, combined with the likelihood of the data, forms an `updated' view of the model, the so-called posterior distribution.
This prior specification can reflect the degree of uncertainty we put into our belief and can be further enhanced by the use of the so called hyperpriors, prior distributions put on the parameters of first order priors.
Moreover Bayesian inference allows for a natural interpretation of the results in terms of probability distributions. % and not somewhat artificial concepts of p-values or critical 
Unfortunately calculating a posterior distribution requires the computation of integrals over the space of all possible parameter values, which are often multi-dimensional.
For a long time, this has limited its use for most but the simplest problems until
\cite{Metropolis1953} presented a Metropolis-Hastings method, %PL: Hasting only came in later, around the 70s, if I remember correctly. He essentially generalized the algorithm. So, referring to the 1953 paper for Metropolis-Hastings  seism wrong to me
Markov chain Monte Carlo (MCMC) algorithm for obtaining a sequence of random samples from a probability distribution for which direct sampling is difficult.%\footnote{This algorithm does not have the most peaceful origins, see the remark.} .

\begin{remark}{Monte Carlo}
Monte Carlo methods are computational algorithms that rely on computer generated pseudo-random numbers to obtain numerical results. 
They were invented in the late 1940's by Stanis\l{}aw Ulam, while he was working on nuclear weapons `project Manhattan'
at the Los Alamos laboratory.  
It was named after the Monte Carlo Casino in Monaco, where Ulam's uncle often gambled.
Monte Carlo was dubbed the most important algorithm of 21st century by the Society for Industrial and Applied Mathematics (SIAM).
\end{remark}

The Bayesian inference problem can be formulated as a sampling problem, as we discuss in Section \ref{sec:bayesian_inference}, reducing to a very simple and effective method, which contributed to the popularity of Bayesian statistics that we see today. 
The availability of programs for Bayesian phylogenetics like MrBayes \citep{Huelsenbeck2001} and BEAST \citep{Drummond2012} further contributed towards the widespread adoption and popularity of these methods in evolutionary biology.

\subsection{Time-Calibrated Trees and Measurably Evolving Populations\label{sub:clocks}}

%Molecular clock
The molecular clock hypothesis (\cite{Zuckerkandl1962}) posits that the rate of evolution of any specified protein remains roughly constant over time and over different lineages.
In particular, Zuckerkandl and Pauling observed that amino acids were replaced at a roughly constant rate in hemoglobins as judged from the fossil record.

% \begin{figure}[H]
% \centering
% \includegraphics[scale=0.3]{molecular_clock} 
% \caption{
% { \footnotesize 
% {\bf Inferred pairwise nucleotide substitutions among 17 mammal species plotted against dates of divergence estimated from fossil records \citep{Wilson1976}.} Linear relationship suggests that pairwise molecular differences are proportional to the time of their separation, and not the morphological traits.
% }% END: footnotesize
% }
% \label{fig:molecular_clock}
% \end{figure}

This hypothesis is compelling because of its simplicity and because it provides a means to date bifurcation events on the tree of life.
A molecular clock assumption can only result in estimates of the proportion of time between two events. 
To estimate concrete dates for phylogenetic nodes, it needs to be calibrated using historical information such as  fossil records.
% MEPs
For rapidly evolving viruses however, it is possible to sample sequences that diverge over an observable timescale. 
As opposed to the ultrametric topology assumed for slowly evolving organism (Figure~\ref{fig:ultrametric}), the dates of the samples themselves can be used to calibrate the molecular clock. %PL: please cite Andrew's TipDate paper here

%GB: I think you should discuss some properties of EIDs at the nucleotide/amino acid/codon level (since that's what we've ended up extending) and link this to what you've done.
%Perhaps you can get some inspiration from the following paper: http://jvi.asm.org/content/84/19/9864.long
%Jeff Thorne is incorporating ''disease properties'' (for example, if they know that a certain non-synonymous mutation can be lethal) into his new codon models, explicitly taking into account likely lethal mutations.

%GB: Can you give a (or some) know example(s) of such substitution or the importance of some particular substitutions?
%How about the fact that a very dangerous virus (which made the news and hence readers will be able to relate to) is only a few mutations away from current bird flu viruses?
%Check this article (the links to the actual Nature and Science papers are at the bottom of the page): http://www.medicalnewstoday.com/articles/246964.php

%---SAMPLING SCHEMES---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.3]{ultrametric} 
\caption{
{ \footnotesize 
%PL: the units of the time scales are hardly visible; please also change 'Heterogenous' to Dated tips and make those labels bigger
{\bf Different tree structures imposed by molecular clock assumptions for organisms evolving at different rates.} The tree on the right represents an evolutionary history of slowly evolving organisms, and has tips that are all aligned at $t=0$. Conversely the tree on the left represents an evolutionary history of rapidly evolving organisms with samples at different points in time, which provide temporal information to calibrate the molecular clock.
}% END: footnotesize
}
\label{fig:ultrametric}
\end{figure}

Time stamped data-sets have also been referred to as heterochronous data and %frequently arise from studies involving viruses which accumulate large amounts of genetic change in a relatively short amount of time.
populations for which such data can be obtained are called measurably evolving populations (MEPs, \cite{Drummond2003}).
This concept has been instrumental for providing estimates of evolutionary rates and for further developments in molecular clock modelling \citep{Drummond2006}.

% relaxed phylogenetics
It was realized relatively early that the molecular clock hypothesis is a simplifying assumption that may be violated by many biological systems.
Several studies demonstrated that the strict molecular clock hypothesis is often rejected for real-world data sets \citep{EyreWalker1997,Andreasen2001}, leading to widespread evidence for heterogenous rates of evolution.
Unfortunately, the alternative `unconstrained' model, which estimates genetic distances for each phylogenetic branch, does not incorporate any relationship between sequence divergence and time and can therefore not be used for divergence time estimation.
%A naive approach to counter these problems is to allow for unconstrained changes in branch substitution rates on the whole tree.
%Unfortunately this does not facilitate the main advantage of the molecular clock hypothesis - the ability to date the speciation events.
This is due to the fact that genetic distances for branches in the unconstrained model are product of the substitution rate and the branch length in time units.
Without any assumption about the rates across branches, both quantities remain confounded (see also Subsection \ref{sub:rates}).

Because both the strict molecular clock model and the unconstrained model are unsatisfactory extremes, many developments aimed at `relaxed molecular clock' models, which allow for a changes in rate over time, but in a constrained manner and therefore still allowing to estimated divergence times.
\citet{Drummond2006}, for example, propose several uncorrelated relaxed clock models in which the substitution rate can vary among the branches of the tree by being drawn independently and identically from some underlying distribution.
Set in a Bayesian framework, this approach allows for a joint reconstruction of the topology and divergence times under less restrictive models.
Among others, a random local clock model approach has also been proposed \citet{Drummond2010} as a more discrete alternative, which allows identifying and quantifying a restricted number of rate changes throughout evolutionary history.
%The RLC model constructs proposal distributions which efficiently propose moves in both the tree space and in the clock rate space at the same time. %(see Subsection~\ref{sub:mcmc})
%Furthermore this approach favours parsimonious models with least parameters, utilizing the fact that for most datasets there exists only a small number of rate changes and that subclades on a tree will share the same rate.
This is achieved through a Bayesian Stochastic Search Variable Selection (BSSVS, \citet{Lemey2009}) procedure that lets the data decide which clock rate changes are needed to explain most rate variation.

\subsection{Processes Giving Rise to Genealogies: Demography and Migration\label{sub:molecular_clock}}

Genealogies are trees that trace the ancestral and descendant relationships among individuals in a population.
The process describing how lineages on a genealogy merge to common ancestor as we go back in time from the present is called the \textit{coalescent} \citep{Kingman1982}.
The simplest formulation of the coalescent process described by John Kingman was derived as an approximation of \textit{Wright-Fisher} \citep{Fisher1930,Wright1931} population model (see Figure~\ref{fig:wright-fisher}).

%---COALESCENT---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.4]{wright-fisher} 
\caption{
{ \footnotesize 
{\bf The Wright-Fisher population model and Kingman's $n$-coalescent} A. A Wright-Fisher population of $N=10$ individuals. The lineages for all individuals sampled from the population coalesce to a common ancestor.
B.  In Kingman's $n$-coalescent, time is continuous and measured from tips towards the root.
}% END: footnotesize
}
\label{fig:wright-fisher}
\end{figure}

Under the Wright-Fisher model the population size $N$ is assumed to be constant through time.
Time (generations) is discrete, i.e. every individual is replaced in every generation, with all individuals having equal chances of reproducing.
To understand how we can describe the coalescent process for a small sample of individuals from such populations as a function of their size, let us start with numbering time (generations), starting from present one as $t_{0},t_{1},t_{2},\ldots,t_{k}$.
The probability that two individuals sampled at the present generation $t_{0}$ share a common ancestor at a preceding generation $t_{1}$ is $\frac{1}{N}$.
The probability that two individuals sampled at $t_{0}$ share an ancestor at $t_{2}$ is the probability that they do not share an ancestor at $t_{1}$, which is $1-\frac{1}{N}$, times the probability of sharing an ancestor at $t_{2}$, which is $\frac{1}{N}$, as the population size remains constant every generation.

%\begin{remark}{Sir John Frank Charles Kingman}
%\begin{figure}[H]
%\centering
%\includegraphics[scale=0.4]{kingman} 
%\caption{
%{ \footnotesize 
%% {\bf Sir John Frank Charles Kingman.} 
%This suave looking gentleman is a British mathematician, with his most notable work on the mathematics of the coalescent, a process fundamental to modern population genetics. 
%Interestingly Kingman never completed his PhD.
%}% END: footnotesize
%}
%\label{fig:kingman}
%\end{figure}
%\end{remark}

Generalizing this, we can write down the probability that a pair of randomly sampled individuals will have their most-recent common ancestor (MRCA) at time $t_{k}$:

\begin{equation}
P\left(t_{k}\right)=\left(\frac{1}{N}\right)\left(1-\frac{1}{N}\right)^{k-1}.
\label{eq:wright-fisher-two-individuals}
\end{equation}

\noindent
Now instead of sampling just two individuals we sample $n,\;2<n\leq N$ individuals, where $n\ll N$.
There are now $\frac{n\left(n-1\right)}{2}$ pairs in $t_{0}$ which can coalesce in $t_{1}$ and each pair has  $\frac{1}{N}$ probability of sharing the same ancestor, therefore:

\begin{equation}
P\left(t_{1}\right)=\frac{n\left(n-1\right)}{2N}.
\label{eq:wright-fisher-n-individuals}
\end{equation}

\noindent
Similarly as in the case of a pair of individuals we can generalize Equation~\ref{eq:wright-fisher-n-individuals} and write down the probability that the sampled pairs of individuals coalesce at $t_{k}$:

\begin{equation}
P\left(t_{k}\right)=\left(\frac{n\left(n-1\right)}{2N}\right)\left(1-\frac{n\left(n-1\right)}{2N}\right)^{k-1}
\label{eq:wright-fisher-n-individuals-general}
\end{equation}

\noindent
Equation~\ref{eq:wright-fisher-n-individuals-general} characterizes a probability mass function (PMF) of a geometric distribution with parameter $p=\frac{n\left(n-1\right)}{2N}$ (see Equation~\ref{eq:geometric_dist}).

\begin{remark}{Geometric distribution}
The geometric distribution returns the probability that the $k$-th trial out of $k$ independent Bernoulli trials, each with success probability $p$, is a successful one:

\begin{equation}
P\left(X=k\right)=p\left(1-p\right)^{k-1},\;0<p\leq1,\; k\in\left\{ 1,2,3,\ldots\right\}.
\label{eq:geometric_dist}
\end{equation}

\noindent
Cumulative distribution function of the geometric distribution:
\begin{equation}
P(X\leq k)=1-\left(1-p\right)^{k-1}.
\label{eq:geometric_cdf}
\end{equation}
\end{remark}

For small values of $p$ the discrete geometric distribution can be approximated by its continuous equivalent, the exponential distribution. 
%An informal argument finds itself in Figure~\ref{fig:exp_approx}.
%
%\begin{figure}[H]
%\centering
%\includegraphics[scale=0.3]{exponential_approx} 
%\caption{
%{ \footnotesize 
%{\bf Approximating geometric distribution by exponential distribution.} 
%Black rectangles denote the discrete geometric distribution, red line is the continuous geometric distribution.
%}% END: footnotesize
%}
%\label{fig:exp_approx}
%\end{figure}
%
This approximation allows us to move into continuous time, and we can write Equation~\ref{eq:wright-fisher-n-individuals-general} as:

\begin{equation}
P\left(t_{k}\right)=\frac{n\left(n-1\right)}{2N}exp\left(-\frac{n\left(n-1\right)}{2N}k\right)
\label{eq:exponential_time}
\end{equation}

In the light of Equation~\ref{eq:exponential_time} we consider the Kingman's coalescent as a continuous time version of the Wright-Fisher model with $n$ sampled individuals.
%Under a molecular clock (see Subsection~\ref{sub:molecular_clock}) mutation rate $r$ and time are confounded.
%%TODO: mutation rate? P(t_k|\theta) ?
%However for MEP populations, were sequences are sampled serially in time (see Figure~\ref{fig:ultrametric}) the two can be disentangled.
\cite{Rodrigo1999} developed an extension of the Kingman's coalescent, the so called \textit{serial coalescent} (s-coalescent), that allows to scale the genealogies in real time units.
Genealogies for a small sample of individuals can be estimated from their genetic data using phylogenetic reconstruction, and the resulting waiting time distributions in the trees can be used to estimate population sizes.
It is important to note, however, that these models make assumptions that are clearly violated in most populations and therefore lead to population size estimates that generally strongly deviate from census population sizes. 
In order to still be able to use those models, a rescaling of the population model has been proposed so that it behaves, with regard to certain properties, as a simple Wright-Fisher model with constant population size.
This rescaling requires moving from census population size to the more abstract concept of `effective population size' or $N_{e}$.
As in changes $N_{e}$ still reflect changes in the census population, the constant population size coalescent model has been extended to account for variable size populations, either in parametric or non-parametric formulations. %PL: add references, e.g. some of the earlier pybus papers fro parametric forms and skyride/skygrid for nonparameterics.
In a Bayesian framework focusing on time-measured trees, such process models find their use as priors on trees.
Coalescent models have also been extended to account for structured populations allowing to estimate migration rates, which naturally brings us to spatial dynamics.  %PL: add references, e.g. Migrate and Lamarc or some of the Hey work for those

% Introduce phylogeography -- discrete ancestral state reconstruction is only one way of doing this (see Bloomquist, Lemey, Suchard)]
How geography impacts genealogies is one the main research questions of phylogeography, a discipline that frequently resorts to ancestral reconstruction to shed light on migration or dispersal dynamics. 
%Another process which shapes the phylodynamics of viral sequences is the process of their migration, which we understand as a dispersal in actual geographical coordinates.  
%Inference methods which make use of both molecular and geographical data have been dubbed as phylogeography as the processes which shape the genome are integrated with the processes of geographical spread.
In that sense the inferred tree does not only provide a record of evolutionary ancestry, but also of the dispersal process that gave rise to the current geographic distribution of taxa.
%
%% snow
%Perhaps a first know study which used epidemiological methods which we today call phylodynamics, is the famous John Snow's work in 1854, tracing the source of a cholera epidemic in the Soho district of London.
%There was no notion of molecular data or even bacteria nor viruses, yet by using a dot map Snow was able to show that almost all the cases in were clustered in a close proximity of a water pump on the Broad Street (see Figure~\ref{fig:snow}).
%Snow was able to persuade local authorities to remove the handle, which contributed to the end of the outbreak.
%It was later found that a nearby cesspool was in fact leaking its content into the public well, therefore transmitting water contaminated with the cholera bacteria.
%
%\begin{figure}[H]
%\centering
%\includegraphics[scale=0.4]{snow_map} 
%\caption{
%{ \footnotesize 
%{\bf Map of the cholera cases in the London epidemic of 1854.} 
%Map shows the clustering of cases around water pump on the Broad Street (currently called Broadwick Street).
%}% END: footnotesize
%}
%\label{fig:snow}
%\end{figure}
%
% stochastic model based
In their opinion piece on current methods in phylogeography, \cite{Bloomquist2010} highlight  fully probabilistic approaches that use Continuous-time Markov chain (CTMC) models to infer transitions between discrete locations as a function of continuous time as one of the phylogeographic inference methods.
Maximum likelihood and closely related Bayesian inference methods assume that both  evolution and geographical dispersal are independent, stochastic processes which unfold on phylogenetic trees.
\cite{Lemey2009} and \cite{Lemey2010} developed a Bayesian framework that allows for a joint reconstruction of spatial diffusion and genetic ancestry, accommodating important sources of uncertainty in estimating spatial and temporal dynamics.
Because they are both modeled using CTMCs, the work presented here on CTMC model extensions finds applications to sequence evolutionary processes as well as these phylogeographic processes.

% population genetic / migration models
%The coalescent which we described at the beginning of this Subsection, and more generally population genetic approaches are another statistical inference method that can be naturally extended for phylogeography.
%These methods assume that trees are shaped by processes such as migration, selection or fluctuations in the population sizes in time.
%Trees proposed by the coalescent are then random draws from a larger, population-level processes, and one can use molecular and other data to get a better resolution of these underlying processes.

\subsection{Bayesian evolutionary inference in BEAST}

The Bayesian evolutionary analysis by sampling trees (BEAST) software \citep{Drummond2012}, to which our research group contributes, integrates many of the advances described in Subsections~\ref{sub:epidemiology}-\ref{sub:molecular_clock}.
BEAST is set in a Bayesian statistical framework, meaning that it combines the data likelihood (see Subsection~\ref{sub:likelihood}) with prior knowledge (see Subsection~\ref{sec:bayesian_inference}), and generates posterior samples using MCMC (see Subsection~\ref{sub:mcmc}).

% other packages
BEAST is not the only software package for Bayesian phylogenetic analysis and can be compared to a number of other packages.
MrBayes \citep{Huelsenbeck2001} is a command-line driven software mainly focusing on phylogenetic inference.
Migrate-n \citep{migrate} and LAMARC \citep{Kuhner2006} on the other hand concentrate on coalescent-driven population genetics.
Like these programs, at the core of BEAST lies a Metropolis-Hasting algorithm (\cite{Metropolis1953, Hasting1970}, see also Subsection~\ref{sub:mcmc}) for constructing Markov chains that converge on the posterior distribution as the equilibrium distribution.
It is precisely why BEAST is not meant for reconstructing a single tree or a point estimate for a parameter of interest, but rather generates samples that can be used to summarize distributions of mutation rates, divergence times, demographic parameters, trait evolution or others.

%beast is cool
When compared to other similar programs, BEAST stands out in the number of models available for performing inference and the flexibility by which they can be combined.
Perhaps the most characteristic point is that BEAST uniquely focuses on sampling rooted, time-measured phylogenies.
This allows for a very natural interpretation of reconstructed trees, and is achieved thanks to the adoption of the molecular clock models, from simple strict clock models, with a uniform rate across the whole time-span of the evolution, to relaxed molecular clock models that model rate variation in various ways \citep{Thorne1998,Yoder2000,Drummond2006,Drummond2010}.
BEAST sports a large number of complementary evolutionary models, which can be combined in a flexible way.
Various substitution models for performing analysis on nucleotide, codon or amino-acid data, accounting for among-site and among-lineage rate variation and partitioned in arbitrary ways, are complemented with demographic tree priors such as coalescent \citep{Drummond2005;Minin2008b} or birth-death priors and trait evolutionary models. %PL: add reference for birth-death

% using beast
%The program takes as input structured XML files which describe the data to be analyzed, the models to be used, the number of samples to collect from the posterior and the output options.
%The output consists of a simple tab-separated files, with each row corresponding to one value sampled from the posterior distribution, for both the parameter values and actual trees.

%gpus
Growing availability of large complex data-sets has led to the term `Big Data'. %being coined.
Large sample sizes, such as full genome sequences or the ones generated with next-gen sequencing techniques are increasingly challenging phylogenetic inference.
Fortunately strategies have been proposed to reduce the time that it takes to fit complex phylogenetic models to the vast amounts of data.
These solutions are mainly focused on parrallellization and on harnessing the powers of modern graphics processing units (GPUs), such as the NVIDIAs Tesla line of GPUs that target the high performance computing market.
BEAGLE (Broad-platform evolutionary analysis generalized likelihood evaluator, \cite{Ayres2012}), is a library that can be used in conjunction with BEAST and provides an API (Application Programming Interface) that massively parallel graphics processors and other multicore devices can exploit to substantially decrease the runtime of phylogenetic analysis \citep{Suchard2009}.
%BEAST readily integrates with BEAGLE and provides a GUI interface to configure the library.

% external software
BEAST is packaged with a number of additional utilities that can either assist in generating the input for BEAST, in analyzing the results or in some other more specific tasks.
\textbf{BEAUti} (Bayesian Evolutionary Analysis Utility) is a  GUI (graphical user-interface) application for generating XML input files for BEAST. 
% It alleviates some of the most cumbersome tasks that would otherwise have to be done manually, thus saving time and simplifying 
It supports the specification of (standard) models, priors, operators, and MCMC chain settings, without having to resort to a manual editing of the XML syntax.
\textbf{LogCombiner} combines log files from multiple runs, that can be than imported into \textbf{Tracer}, a graphical tool for the analysis and visualization of the output of BEAST as well as other common MCMC packages.
It allows for a quick inspection of mixing, convergence, and summary statistics of the parameters of interest.
\textbf{TreeAnnotator} summarizes the tree log files and generally represents the evolutionary history using a maximum clade credibility (MCC) tree; each node or branch in this tree can be annotated with external information, such as posterior probabilities and maximum a posteriori ancestral traits.
The tree and its annotated information can subsequently be visualized using \textbf{FigTree} (\url{http://tree.bio.ed.ac.uk/software/figtree/}). % or if nodes of the trees are annotated with spatial information they can be displayes in geographical coordinates using \textbf{SPREAD}. %PL we need to introduce spread first, which is a chapter in this thesis :-), same for piBUSS
%Finally \textbf{$\pi$BUSS} is a software for sequence simulation, capable of generating XML files understood by BEAST for combined simulation-inference analysis as well as sequences in formats taht can be imported into any other phylogenetic software.
%
In conclusion, BEAST is a flexible package for statistical pylogenetics and hypothesis testing.
The software is continuously being improved and new models are being developed, but the development has recently been split into a continuing BEAST1 branch and a new BEAST2 branch. %PL: add BEAST2 reference.
Since the work presented in this thesis is implemented in BEAST1, all references to BEAST will implicitly refer to this development branch.
%In addition to a large array of available models, it is actively developed with new models and enhancements included with every new stable release.

\subsection{Objectives: model extension, simulation and visualisation\label{sub:objectives}}

% models
\paragraph{}
Phylogenetic analyses employing stochastic processes make strong restrictive assumptions regarding the underlying substitution process.
These assumptions are mainly enforced to reduce the number of free parameters of the models and to ease computational and mathematical tractability.
Unfortunately many of these assumptions hamper the application of the methods to complex, real-world data-sets.
Relaxing some of these assumptions and development of more biologically plausible models, able to grasp the complex nature of phylodynamic dispersal processes is an important aim of this thesis \citep{Bielejec2014a}.
%Another goal is to make the general applicability suitable for epidemiological problems as well as beyond them, with potential to turn the findings into useful predictions of the emergence of infectious diseases, e.g. by identifying the sources and sinks of the viral spread.
% geographical dispersal

% simulation
\paragraph{}
The development of novel phylogenetic inference methods brings about the need to compare estimator performance and to characterize strengths and weaknesses of different approaches (e.g. \cite{Arenas2012}, \cite{Hoban2011}).
Whereas the true underlying evolutionary relationships between biological sequences are generally unknown, Monte Carlo simulations allow generating test scenarios while controlling for the evolutionary history as well as the tempo and mode of evolution. 
Part of the research presented in this thesis is devoted to the development of a flexible software for sequence simulation, that integrates both the tree-generative processes, as well as models responsible for shaping the molecular histories and geographical dispersal \citep{Bielejec2014a}.

% visualisation
\paragraph{}
Bayesian phylogenetic inference inevitably results in complex, multi-dimensional output.
Adding the geography to the phylogenetic process further adds to this complexity.
In particular, Bayesian phylogeography  result in posterior distributions of trees, with each tree branch annotated with estimated node locations and other traits of interest.
%This represents a major challenge for visualization  the phylogenetic inference.
Clear and intuitive visualizations are an important aspect of every analysis \citep{Hadley2010}, but this is far from trivial for the estimates resulting from Bayesian analysis of sequences and traits.

Spatial phylogenetic projections have already been utilized in the field \citep{Kidd2006,Parks2009}, yet most of these applications remain limited to mapping phylogenetic tip taxa to their geographical coordinates, without providing robust statistical estimates of the geographic locations at the ancestral nodes nor the measures of uncertainty in the inference.
Fortunately the advent of powerful and flexible Geographic Information Systems (GIS) has fostered an interesting possibility to visualize both the spatial and temporal information resulting from viral phylogenetic analysis.
Presenting readily interpretable visual summaries of the inferences in e communicated to the researchers across different fields is one of the more important aspects of this thesis.
%This goal has been materialized in the development of a user-friendly application, that produces compelling and informative statistical visualizations of viral spread in time and space \citep{Bielejec2011}.

\section{Evolution as a stochastic process}

In order to provide a formal basis for the work in the research chapters of this thesis, this chapter proceeds with a description of general stochastic models of character substitution processes in phylogenies. 
Molecular phylogenetic modeling typically starts with an assumption that the observed sequence data has been generated by some stochastic process, which emits characters over time.
We can think of these stochastic processes as functions of time, which is regarded as a deterministic argument, whose values are random (non-deterministic) variables. 
It is thus a statistical model of a random development, or evolution, in time.
Phylogenetic inference often resorts to processes where time is considered to be continuous and the indexed set of outcomes, called the state space, is countable and finite (discrete).
% The stochastic process $X$ is a collection $\left\{ X(t):\; t\in T\right\} $, where each drawn value $X(t)$ is a random variable drawn from state space $\mathcal{E}$, with $K$ possible values, indexed by orderer time $T$.
With that in mind we can provide a formal definition:

\begin{definition} 
The stochastic process $X$ is a collection $\left\{ X(t):\; t\in T\right\} $, where each drawn value $X(t)$ is a random variable drawn from state space $\mathcal{E}$, with $K$ possible values, indexed by orderer time $T$.
\label{def:stochasticProc}
\end{definition}

In the simplest case, for which the unit of evolution is a single nucleotide, we have $\mathcal{E}=\left\{ A,C,G,T\right\}$ and $K=4$.
%GB: perhaps point out that they're in fact not independent?
The particular sites of the sequence alignment are then considered to be stochastically independent realizations of the process $X$, leading to the observed data (see Figure~\ref{fig:alignment}).

\begin{figure}[H]
\centering
\begingroup
\everymath{\displaystyle}
{\Large
\begin{displaymath} % 
\xymatrix{
& \color{gray2}A \ar[dl] \ar[drdr]^{ \color{gray2}\begin{array}{c}
A\rightarrow G\\
G\rightarrow C
\end{array}} & & \\
T & & & \\
& & & C \\
}
%     \xymatrix@R+1pc{ 
% \color{gray2}A \ar[r] & \color{gray2}T \ar[r] & A \\
% \color{gray2}A \ar[r] & \color{gray2}T \ar[r] & G \\
% \color{gray2}A  \ar[rr]_{ \color{gray2}\begin{array}{c}
% A\rightarrow G\\
% G\rightarrow C
% \end{array}} &&  C
%     } % 
\end{displaymath}
}% END: Large
\endgroup
\caption{
{ \footnotesize 
{\bf Evolution at a single site.} Over time characters at the site change, leading to the observed data (black). 
%Some of the mutations may be unobserved. %PL: All of the mutations are unobserved
Each site evolves independently.
The longer the evolutionary time, the higher the probability that multiple substitutions occur on the same site.
}% END: footnotesize
}
\label{fig:alignment}
\end{figure}

As illustrated in Figure~\ref{fig:alignment}, multiple substitutions (multiple hits) may occur, which complicates the estimation of the underlying evolutionary distances from the observed data.
Specifically, the number of observed differences in the sampled genetic data will be an underestimate the underlying genetic distance (in Figure~\ref{fig:alignment}, three substitutions underlie a single difference in the observed states).
In order to accurately describe these processes one needs a model of evolution that can account for those multiple hits.

\subsection{The Poisson process\label{sub:poisson}}

The Poisson process is an example of a continuous-time stochastic process that is frequently used to describe counts of independent, rare events, occurring with some intensity (or rate) $r$, that remains constant over time. 
Examples of such a process are the requests for a single document on a web-server, or the goals scored in a football game \footnote{Unless it's Real - Barca, than the events are not so rare.}. %PL or Brazil - Germany...
%The state space in this case is obviously $\mathcal{E} = \left\{ 0,1,2, \dots \right\}$.
The Poisson process is equally suitable to model the number and waiting times of DNA mutation of sequence substitution events. 
We can formally define a Poisson process as following:

\begin{definition} 
A Poisson process with rate $r>0$ is a continuous-time counting process $\left\{ N(t):\; t\geq0\right\}$ such that:

\begin{itemize}
\item $N(0)=0$.
\item The increments are independent i.e. if $(t_{1},t_{2}]\cap(t_{3},t_{4}]=\emptyset$ then $N(t_2)-N(t_1)$ and $N(t_4)-N(t_3)$ are independent and stationary (i.e. the probability distribution of the number of occurrences counted in any time interval only depends on the length of that interval).
\item The number of events in any interval of length $t$ is $\text{Poisson}(r t)$.
\end{itemize}

\label{def:poisson}
\end{definition}

%PL: I commented the figure out for now because you are representing coalescent waiting times and not mutational waiting times; just show a number of Poisson mutation trajectories with more mutations overlaid or the idea that we discussed with A,C,G,T states on the Y-axis

%Figure~\ref{fig:poisson} shows a sample of different trajectories of such a process.

%
%% TODO
%
%%PL: I don't understand the figure composition: upper part OK, but if you use color as you do now, different realizations can be plotted on the same panel. However, I would keep the printing issues for color figures  to a minimum (only use colors when you really need it -- it requires separate paper printing and merging with the rest, and is bloody expensive), so separate gray-scale panels can work. bottom part: is what I do not understand: should the paths assemble into a tree somehow? The discussion of the process makes no relationship to a tree just yet -- you do that later on. And if you do intend to relate them to a tree, I do not understand how the mutational process for a site only relates to part of the tree?
%%FB: I wanted to show that different paths (diff branch lengths) are independent relaisations, and waiting time is exponential (similar to Marcs 'travelling particle' description from one of your papers). This is in greyscale now, but if you feel lik eit doesn't conect then we can just comment this figure out.
%\begin{figure}[H]
%\centering
%\includegraphics[scale=0.4]{poisson}
%\caption{
%{ \footnotesize 
%{\bf Independent realizations of the Poisson process with intensity $r=1$.}
%Times between mutations (the branch lengths) are \emph{iid} exponentially distributed random variables with mean $1/r$.
%} % END: footnotesize
%}
%\label{fig:poisson}
%\end{figure}

Let's assume that evolution at a single site is governed by a Poisson process $N$ with some intensity $r$.
Definition~\ref{def:poisson} implies that under this model the number of mutations $N(t+u)-N(u)$ in some time interval $(u,t+u]$ of length $t$ follows a Poisson distribution with expected number of events $r t$.
Thus we can write:

\begin{equation}
P\left\{ N(t+u)-N(u)=k\right\} =P\left\{ N(t)=k\right\}=\frac{e^{-r t}(r t)^{k}}{k!},\; k=0,1,\ldots
\label{eq:poissonDist} 
\end{equation}

In the Poisson process the time between any two mutations is exponentially distributed with parameter $r$, thus the average waiting time between the occurrence of events is $1/r$.
To show this important property we can consider $T_1,T_2,\ldots$ to be the inter-arrival times, where $T_k$ is the waiting time between the $(k-1)\text{'st}$ and $k\text{'th}$ mutation. 
Obviously, the number of mutations before some time $t>0$ is less than some integer $k>0$ if and only if the waiting time $T_k$ until the $k\text{'th}$ mutation is larger than $t$:

\begin{equation}
P\left(N(t)<k\right)=P\left(T_{k}>t\right).
\end{equation}

\noindent
For the first waiting time $T_1$ we have:

\begin{equation}
1-P\left(T_{1}>t\right)=1-P\left(N(t)<1\right)=1-P\left(N(t)=0\right).
\end{equation}

\noindent
From Equation~(\ref{eq:poissonDist}) we have that:

\begin{equation}
F_{T_{1}}(t)=1-P\left(T_{1}>t\right)=1-e^{-r t},
\end{equation}

\noindent
therefore $T_1$ follows the exponential distribution with parameter $r$.
For the second waiting time $T_2$ and some arbitrary $u>0$ and $t>0$, the independent increments property implies that:

\begin{eqnarray}
F_{T_{2}}(t) &=& 1-P\left(T_{2}>t|T_{1}=u\right) \\ \nonumber
& = & 1-P\left(\text{no mutations in }(u,t+u]|T_{1}=u\right) \\ \nonumber 
& = & 1-P\left(N(t)=0\right)=1-e^{-r t}.
\end{eqnarray}

\noindent
Therefore $T_2$ is independent of $T_1$ and $\text{Exponential}(r)$. 
Similarly we can show that $T_3$ is independent of $T_1$ and $T_2$ and by repeating the same argument $T_1,T_2,\ldots$ are independent and identically distributed (i.i.d.) $\text{exponential}(r)$ random variables.

% TODO: memoryless property does not depend at all, vide poiss)
% TODO: markovian (depends only a little)
% * difference? Memoryless is more

\subsection{Instantaneous rates of mutation\label{sub:rates}}

Within a phylogenetic framework the interest typically lies not only in modeling the number of events (mutations), but also the actual probabilities of changing states at the particular site in the alignment. 
We therefore need to generalize the process to a finite number of discrete states.
Let $\mathbf{R}$ denote a $K \times K$ matrix filled with probabilities of exchanging states between any pair $i,j\in \mathcal{E}$.
To express the probability of going from state $i$ to $j$ in some time interval $t$, we multiply the probability of the state change by the probability of having exactly $k$ events leading to the observed change, and integrate over all possible values of $k$.
This allows us to accommodate all possible paths that evolution might have taken, and to account for multiple hits that might have occurred at the site, as illustrated in Figure~\ref{fig:alignment}.
From Equation~(\ref{eq:poissonDist}) we know the probability of having $k$ mutations in time interval $t$, thus:

\begin{equation}
\left\{ P(t)\right\} _{ij}=\underset{k=0}{\overset{\infty}{\sum}}\mathbf{R}^{k}\frac{(r t)^{k}}{k!}e^{-r t}=\underset{k=0}{\overset{\infty}{\sum}}\frac{(\mathbf{R}r t)^{k}}{k!}e^{-r t},
\label{eq:matrixExp1}
\end{equation}

\noindent
where $\left\{ P(t)\right\} _{ij}$ denotes the $ij\text{-th}$ element of the transition probability matrix $\mathbf{P}$ over time $t \geq 0$.
Let us denote the matrix $\mathbf{R}-\mathbf{I}$, where $\mathbf{I}$ is a $K \times K$ identity matrix, by $\mathbf{Q}$.
Using the power series definition of the transition probability matrix $\exp(\mathbf{R})=\underset{k=0}{\overset{\infty}{\sum}}\mathbf{R}^{k}/k!$ we can write Equation~(\ref{eq:matrixExp1}) as:

\begin{equation}
\left\{ P(t)\right\} _{ij}=e^{\mathbf{R}r t}e^{-r t} =e^{(\mathbf{R}-\mathbf{I})r t}=e^{\mathbf{Q}r t} 
\label{eq:matrixExp2}
\end{equation}

\noindent
A single entry $q_{ij} \ge 0$ for $i \neq j$ in matrix $\mathbf{Q}$ quantifies the probability of changing from state $i$ to $j$ in an infinitely small amount of time, therefore we will refer to $\mathbf{Q}$ as the \emph{instantaneous rate matrix}.

\subsection{Computing the matrix exponent \label{sub:exponentiation}}

In order to arrive at the finite-time transition probabilities matrix $\mathbf{P}(t)$ for some time $t\geq0$, one needs to compute the matrix exponential in Equation~\ref{eq:matrixExp2}.
\cite{Moler1978} provide an excellent review of numerical methods to compute the matrix exponent. 
Here, we will highlight the method utilized by most phylogenetic software, including BEAGLE \citep{Ayres2012} and BEAST \citep{Drummond2012}, that uses the singular value decomposition (SVD) of the rate matrix $\mathbf{Q}$.

Let us assume $\mathbf{Q}$ has $K$ linearly independent eigenvectors $\mathbf{v}_{i} \; (i=1,\ldots,K)$ with corresponding eigenvalues $\lambda_{i}\;(i=1,\ldots,K)$:

\begin{equation}
\mathbf{Q}\times\mathbf{V}=\lambda\times\mathbf{V}
\label{eq:eigenvaluesEigenvectors}
\end{equation}

\noindent
This condition is sufficient for $\mathbf{Q}$ to be diagonizable, thus its Jordan form is:

\begin{equation}
\mathbf{V}\times\mathbf{Q}\times\mathbf{V}^{-1}=\left[\begin{array}{cccc}
\lambda_{1} & 0 & \cdots & 0\\
0 & \lambda_{2} & \ddots & \vdots\\
\vdots & \ddots & \ddots & 0\\
0 & \cdots & 0 & \lambda_{K}
\end{array}\right]=\text{diag}(\lambda_{1},\ldots,\lambda_{K})=\mathbf{D} ,
\label{eq:diagonalization}
\end{equation}

\noindent where $\mathbf{V}$ is a matrix composed of eigenvectors of $\mathbf{Q}$. 
Multiplying both sides of Equation~(\ref{eq:diagonalization}) with the edge length $t$ and scaling by the appropriate rate scalar $r$, we can write $\mathbf{Q}$ in the following form:

\begin{equation}
rt\mathbf{Q}=\mathbf{V}\times\text{diag}(rt\lambda_{1},\ldots,rt\lambda_{K})\times\mathbf{V}^{-1} .
\end{equation}

\noindent Finally, using the power series definition of the transition probability matrix, we obtain:

\begin{eqnarray}
\mathbf{P}(t) &=& \exp\left(rt\mathbf{Q}\right)  =  \underset{k=0}{\overset{\infty}{\sum}}\left(rt\mathbf{Q}\right)^{k}/k! \\ \nonumber
 &=&  \underset{k=0}{\overset{\infty}{\sum}}\left(\mathbf{V}\times\mathbf{D}\times\mathbf{V}^{-1}\right)^{k}/k! \\ \nonumber
 &=& \mathbf{V}\times\left(\underset{k=0}{\overset{\infty}{\sum}}\mathbf{D}^{k}/k!\right)\times\mathbf{V}^{-1} = \mathbf{V}\times\text{diag}(e^{rt\lambda_{1}},\ldots,e^{rt\lambda_{K}})\times\mathbf{V}^{-1} ,
% \mathbf{P}(t) & = & \exp\left(rt\mathbf{Q}\right)  =  \underset{k=0}{\overset{\infty}{\sum}}\left(rt\mathbf{Q}\right)^{k}/k! 
%  =  \underset{k=0}{\overset{\infty}{\sum}}\left(\mathbf{V}\times\mathbf{D}\times\mathbf{V}^{-1}\right)^{k}/k! \\ \nonumber
% & = & \mathbf{V}\times\left(\underset{k=0}{\overset{\infty}{\sum}}\mathbf{D}^{k}/k!\right)\times\mathbf{V}^{-1} = \mathbf{V}\times\text{diag}(e^{rt\lambda_{1}},\ldots,e^{rt\lambda_{K}})\times\mathbf{V}^{-1} ,
\label{eq:eigen_decomposition}
\end{eqnarray}

\noindent 
where the last transition is due to the fact that the exponential of the diagonal matrix is obtained by exponentiating every entry on the main diagonal.
We should note here that the repeated evaluations of $\exp\left(rt\mathbf{Q}\right)$ for different times $t$ are all based on the same SVD of the matrix $\mathbf{Q}$. 
One needs only to scale and exponentiate the eigenvalues to calculate the diagonal matrix $D$ and perform two matrix multiplications, thus reducing all calculations to a simple and effective algorithm.

\subsection{Markov chain models of sequence substitution\label{sub:subst_models}}

% TODO markov property is inherited from Poisson process
Drawing realizations with probabilities defined by $\mathbf{P}$  gives rise to a continuous-time stochastic process $\left\{ X(t):\; t\geq0\right\}$ satisfying the Markov property, such that for every $n\geq 0$, given the time points $0\leq t_{0}\leq t_{1}<\ldots<t_{n}\leq t_{n+1}$ and discrete states $i_{0},i_{1}, \ldots, i_{n},i_{n+1}$, it holds that:

\begin{eqnarray}
P\left\{ X(t_{n+1})=i_{n+1}\mid X(t_{n})=i_{n},\ldots, X(t_{0})=i_{0}\right\} & = & \\ \nonumber
P\left\{ X(t_{n+1})=i_{n+1}\mid X(t_{n})=i_{n}\right\} .
\label{eq:markov}
\end{eqnarray}

The elements of $\mathbf{P}$ quantify the finite-time transition probabilities between the $K$ discrete state-space elements.
This describes a Continuous-time Markov chain (CTMC), which  we will use to model the substitution process at a single site of an alignment.
Every CTMC is completely characterized by its rate matrix $\mathbf{Q}$ and the stochastic matrix $\mathbf{P}(t),\ t\geq0$ is obtained via matrix exponentiation as described in the previous subsection (Subsection~\ref{sub:exponentiation}).

%PL: Not sure if you want to do a color page print for this figure; it could be black/white

%---SUBSTITUTION MODEL---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{substitution} 
\caption{
{ \footnotesize 
{\bf  Graphical representation of  a nucleotide substitution model.} A Markov substitution model with 6 parameters describing the symmetric transition rates.
} % END: footnotesize
}
\label{fig:substitution}
\end{figure}

Figure~\ref{fig:substitution} depicts an example of such a CTMC substitution model, governing the substitution rates between four discrete nucleotides, where all the pairs communicate (all states have a non-zero transitional probability of exiting from, i.e. all states are non-absorbing).
Let us denote a transition probability between two states $i$ and $j$ over time $u$ to $t+u$ by:

\begin{equation}
p_{ij}\left(u,t+u\right)=P\left\{ X(t+u)=j\mid X(u)=i\right\} .
\end{equation}

In the phylogenetic setting, researchers often constrain the CTMC processes, mostly for computational and mathematical convenience.
A frequently applied restriction is to assume time-homogeneity of the substitution process, which states that the evolution does not change in pattern over time, and therefore transition probabilities depend only on the time difference $t$ between $u$ and $t + u$:

\begin{equation}
p_{ij}\left(t\right) = p_{ij}\left(0,t\right) = p_{ij}\left(u,t+u\right).
\label{eq:time_homogeneity}
\end{equation}

Unfortunately this assumption does not hold for many real-world data-sets and precludes the identification of biologically interesting patterns in the evolutionary process. %, as the temporal dynamics clearly do shift in time, often in a seasonal, observable manner \citep{Bahl2011}.
%PL: deleted the above subsentence as it is too specific to phylogeography
One of the goals of the research presented in this thesis is to efficiently confront situations in which this assumption does not hold.
% can grasp such situations.

Substitution models are further constructed to be ergodic, meaning that there exist positive values $\pi_{1},\pi_{2},\ldots,\pi_{K}$ such that:

\begin{equation}
\forall i,j\in \mathcal{E}  \quad\underset{t\rightarrow\infty}{lim}p_{ij}(t)=\pi_{j}.
\label{eq:ergodicity}
\end{equation}

% $\mathbf{\Pi}$

As $t$ goes to infinity the probability that the site is in some state $j$ is non-zero and does not depend on the starting state $i$.
This implies that the chain will always display the same long-term behavior, which is why $\mathbf{\Pi}=\{\pi_{i},\ i\in\mathcal{E}\}$ is called a \emph{stationary distribution} (also referred to as the \emph{equilibrium distribution} or \emph{equilibrium frequencies}).
We can look at the elements of $\mathbf{\Pi}$ as the proportion of time spent in state $j\in\mathcal{E}$ after the Markov chain has run for infinitely long time, that does not depend upon the initial state $i$ of the process.
In other words if we sample the initial state from the stationary distribution and then let the process run for some sufficiently long time $t$, then the distribution of the final state will equal the stationary distribution.

Another commonly applied constraint is that of \emph{time-reversibility}, meaning that the model does not care which character is the ancestor and which is the descendant so long as all other parameters are held constant.
For all $i,j\in \mathcal{E}$ and $t\geq 0$ we have:

\begin{equation}
\pi_{i}p_{ij}(t)=\pi_{j}p_{ji}(t).
\label{eq:time_reversibility1}
\end{equation}

\noindent
Condition~\ref{eq:time_reversibility1} is also referred to as the \emph{detailed balance} equation, and posits that the probability of sampling $i$ from the stationary distribution and then going to $j$ over some finite-time $t$ is the same as sampling $j$ and going to $i$. 
% By the definition of $\mathbf{\Pi}$ and $\mathbf{P}$ from (\ref{eq:time_reversibility1}) we have:
% 
% \begin{equation}
% \pi_{i}q_{ij}=\pi_{j}q_{ji}\mathbf{Q}.
% \label{eq:time_reversibility2}
% \end{equation}
% 
% Definition (\ref{eq:time_reversibility1}) and (\ref{eq:time_reversibility2}) are equivalent.
This condition is assumed for strictly computational reasons, i.e. time-reversibility makes it easier to diagonalize $\mathbf{Q}$, which - as we have described in 
Subsection~\ref{sub:exponentiation} - is a strategy frequently employed in phylogenetic software to compute the exponentials of rate matrices.
Time reversibility also simplifies the computation of the likelihood on a tree, since it makes the computation independent of the position of the root, which we will discuss in Subsection~\ref{sub:likelihood}.

%TODO: JK69 as example
\subsection{A simple example: the Jukes and Cantor substitution model\label{sub:jc69}}

In this subsection we will discuss the simplest model of nucleotide substitution, i.e. the JC69 model \citep{Jukes1969}.
The JC69 model assumes that the instantaneous rates of change between every two nucleotides are the same.
We also require that the rate matrix $Q$ characterizing the model is a stochastic matrix (each row sums to 0).
The infinitestimal rate matrix of the JC69 model is thus given by:

\begin{equation}
\mathbf{Q}=\left[\begin{array}{cccc}
-3r & r & r & r\\
r & -3r & r & r\\
r & r & -3r & r\\
r & r & r & -3r.
\end{array}\right]
\label{eq:jc69}
\end{equation}

\noindent
The finite-time transition probability matrix $\mathbf{P}(t)$ for time $t>0$ is calculated using matrix exponentiation  (see Subsection~\ref{sub:exponentiation}):

\begin{eqnarray}
\mathbf{P}\left(t\right) &=& \left[\begin{array}{cccc}
p_{0}(t) & p_{1}(t) & \cdots & p_{1}(t)\\
p_{1}(t) & p_{0}(t) & \ddots & \vdots\\
\vdots & \ddots & p_{0}(t) & p_{1}(t)\\
p_{1}(t) & \cdots & p_{1}(t) & p_{0}(t)
\end{array}\right], \\ \nonumber
&\text{ where }& \ensuremath{\begin{cases}
p_{0}(t)=\frac{1}{4}+\frac{3}{4}e^{-4rt}\\
p_{1}(t)=\frac{1}{4}-\frac{1}{4}e^{-4rt}
\end{cases}}.
\label{eq:jc69Finite}
\end{eqnarray}

\noindent
A few things are worth mentioning about Equation~(\ref{eq:jc69Finite}).
If we let $t\rightarrow \infty$ we can see that $\forall i,j\; p_{ij}(t)=1/4$, which means that the equilibrium distribution for the JC69 model is:

\begin{equation}
\mathbf{\Pi}=\left\{ \pi_{T}=1/4,\pi_{C}=1/4,\pi_{A}=1/4,\pi_{G}=1/4\right\}
\label{eq:jc69Steady}
\end{equation}

\noindent
Second, the rate and time in Equation~(\ref{eq:jc69Finite}) occur as a product, which means that rate and time are confounded.
If we were to simulate two sequences, using the algorithm listed in (\ref{alg:simulation}) and starting the random number generators form the same seed, they will look the same for every $r$ and $t$ as long as their product $rt$ is the same.
This is true for most of the substitution models and implies that typical phylogenetic estimation focuses on inferring the product parameter $\theta=rt$, or genetic distance.
%PL: perhaps mention that the section on molecular clocks (cfr supra) discusses the conditions for which we can start separating time and rate

\subsection{Phylogenetic trees}

%PL: According to the extension I recommend for the general intro, a tree should already be introduced properly there, and this should from the biological perspective but can also include this more formal description..

A phylogenetic tree, or phylogeny for short, depicts the genealogical relationship between character sequences.
Mathematically, a phylogeny $\mathbf{F}$ represents a directed, acyclic graph with sets of \emph{nodes} connected by \emph{branches} (e.g., Figure~\ref{fig:treeconcept}).
Data is observed only at the external nodes (\emph{tips}), while internal nodes represent extinct ancestors for which data is usually missing.
The ancestor of all observed samples is called the \emph{root} of the tree and it's height is referred to as the time to the most recent common ancestor (TMRCA).

%---TREE CONCEPT---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{treeconcept} 
\caption{
{ \footnotesize 
{\bf A phylogenetic tree showing evolutionary relationship.} Character sequences are observed at the tips of the tree (black dots); sequence characters at the internal nodes are treated as missing data. 
Branch lengths are estimated in units of genetic distance, or in this case, real time.
} % END: footnotesize
}
\label{fig:treeconcept}
\end{figure}

Within our framework we focus on trees which are rooted and time-measured, i.e. their branch lengths are scaled in actual time units.
A rooted binary tree with $n$ leaves has $2n-2$ branches and $n-1$ internal nodes.
The branching pattern of the tree is called the \emph{topology}.

\subsection{Sequence evolution on a tree\label{sub:evolutionOnTree}}

In this section we proceed by describing the connection between the CTMC models of character substitution discussed above and phylogenetic histories.
We keep the focus on a single site of the alignment, and due to the assumption of independence the probability of observing an alignment is simply the product of probabilities for observing the individual sites.

To gain a better understanding of the process shaping the data observed at the tips of a phylogenetic tree let us consider the problem of simulating sequence evolution on a tree according to a model with rate matrix $\mathbf{Q}$.
We start with a first character $i$ sampled at the root of the tree $x_0$ from the stationary distribution $\mathbf{\Pi}$.
Let us assume that two direct descendent nodes are $x_j$ and $x_k$.
The character sampled at the root waits an exponentially distributed amount of time $t\geq0$ (see Figure~\ref{fig:poisson}) before moving to one of the descendent or child nodes.
Such traversal is said to be following the pre-order, i.e. parental nodes are visited before child nodes.   
The process then randomly decides by a coin flip to which of the two descendent nodes to move next (assuming that the visited node is not a tip).
Let us assume that the process moved to $x_j$.
The state $j$ for the node $x_j$ is sampled conditional on the value $i$ already sampled at the root using the branch length expressed in real time $t$ and the vector probabilities in the row $i$ of the finite-time probabilities matrix $\mathbf{P}(t)$, calculated using SVD of the rate matrix $\mathbf{Q}$ that characterizes the CTMC.

\begin{algorithm}[H]
\centering
\begin{algorithmic}[1]
% \footnotesize{
%
\State $node \gets getRoot\left(\right)$
%
\State $state \gets getState\left(node\right)$
%
\Repeat
%
\If{$\left(hasChildren\left(node\right)\right)$}
%
\ForAll{children}
%
\State $t \gets getDistanceToParent\left(child\right);$
%
\State $ \mathbf{P}\left(t\right) \gets e^{\mathbf{Q}t};$
%
\State $state \gets sample\left(\mathbf{P}\left[ state, \right]\right);$
%
\EndFor
%
\Else \Comment{this is tip node}
%
\State $node \gets getParent\left(node\right);$
%
\EndIf
%
\Until{$\left(\text{simulated for all nodes}\right)$}
% }
\end{algorithmic}
\caption{
{ \footnotesize 
%PL: Perhaps add: Pseudo code for simulating...
{\bf Simulation an evolutionary process along a phylogeny.} 
When a child node is visited, the state is sampled with conditional probabilities of changing to state $j$ given state $i$ at the parental node.
}% END: footnotesize
}
\label{alg:simulation}
\end{algorithm}

This illustrates the Markov property of a CTMC, as defined in Equation~(\ref{eq:markov}).
The sampling is then repeated on the node $x_j$, before recursively proceeding to the other trio of nodes.
The states sampled at the tips of the tree constitute the character sequence for one site of the alignment.
In this light the CTMC process can be viewed as a continuous-time random-walk, unfolding on the topology $\mathbf{F}$, with its behaviour described by a rate-matrix $\mathbf{Q}$.
The complete algorithm is listed in Algorithm~\ref{alg:simulation}.

\section{Likelihood of sequence evolution}

\subsection{Calculation of the likelihood on a tree\label{sub:likelihood}}

Likelihood is defined as a conditional probability of observing the data given the parameters, as a function of those parameters.
%PL: mention perhaps Fisher, Edwards..?
Phylogenetic inference considers discrete molecular sequence data in the form of aligned matrix $\mathbf{X}$ of size $n \times l$, where $x_{jh}$ is the $h$-th character in $j$-th sequence.
Single column $\mathbf{x}_{h}$ of the data matrix constitutes a single site.
Again, the standard assumption of independence posits that different sites evolve independently and that once two branches split at a node, the evolution on those branches is also independent.

In Subsection~\ref{sub:evolutionOnTree} we described how evolution according to some process $\mathbf{Q}$ gives rise to the observed data $\mathbf{X}$.
In the typical situation we start with just the observed data and aim the inference efforts at modeling the process that generated the alignment.
We can therefore denote the likelihood simply as:

\begin{equation}
L\left(\mathbf{\Theta}|\mathbf{X}\right)=P\left(\mathbf{\Theta}|\mathbf{X}\right), 
\label{eq:likelihood}
\end{equation}

\noindent
where $\mathbf{\Theta}$ represents the tree topology, the branch lengths, the parameters of the substitution model and other parameters of interest that we aim at estimating.
This quantity expresses how likely it is to observe the data given a specific value of the parameter of interest $\mathbf{\Theta}$.
By finding the value of $\mathbf{\Theta}$ that maximizes (\ref{eq:likelihood}) we capture a model with the best possible fit to the data.

In this thesis we will not discuss the maximum likelihood estimation itself, %PL: good, who wants to consider the data to be fixed anyway..
rather focus on closely related 
%PL: I would add: ".., but philosophically very different,"
methods of Bayesian inference.
The likelihood concept is also required in Bayesian inference and enters the framework as the data component.
We will now discuss how to calculate the likelihood of observing a particular realization of character sequences at the tips of a tree, given a substitution model with fixed parameters.

%---LIKELIHOOD---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{likelihood} 
\caption{
{ \footnotesize 
{\bf  A hypothetical tree for five taxa.} Character data for single site is observed at the tips of the tree.
} % END: footnotesize
}
\label{fig:LIKELIHOOD}
\end{figure}

%PL: I could recommend to align the char's, increase their size, relate them to the tips with horizontal dashed lines, but just increasing their size may do
% also, why not label the external nodes with x1-to x5 as you mention in the text, consistent with the internal models?
% the i indices do not correspond between the nodes and branches for x7 and x8

To calculate the likelihood means in a sense to reverse-engineer the evolution as it happened.
Figure~\ref{fig:LIKELIHOOD} presents single site with data $TCACC$ observed at the tips of the tree.
External nodes are numbered $x_{1}, \ldots x_{5}$, internal nodes $x_{6}, \ldots x_{8}$ and the root is placed at the node $x_{0}$.
The length of each branch leading to node $i$ is denoted as $t_{i}$. %PL: the i indices do not correspond between the nodes and branches for x7 and x8
Using the methods presented in Subsection~\ref{sub:exponentiation} we calculate the finite-time transition probabilities $p_{ij}(t_{i})$ between all pairs of characters in the alignment.
To calculate the probability at a single site, we need to sum over all possible combinations that could give rise to the observed character data at that site, in this case:

\begin{eqnarray}
\underset{x_{0}\in \mathcal{E}}{\sum}\;\underset{x_{6}\in \mathcal{E}}{\sum}\;\underset{x_{7}\in \mathcal{E}}{\sum}\;\underset{x_{8}\in \mathcal{E}}{\sum}[\pi_{x_{0}}\cdot p_{x_{0}x_{8}}(t_{7})\cdot p_{x_{0}x_{7}}(t_{8})\cdot p_{x_{8}x_{6}}(t_{6})\cdot p_{x_{8}A}(t_{3}) \cdot \\ \nonumber
p_{x_{7}C}(t_{4})\cdot p_{x_{7}C}(t_{5})\cdot p_{x_{6}T}(t_{1})\cdot p_{x_{6}C}(t_{2})]
\label{eq:likelihoodNaive}
\end{eqnarray}

We can break down the terms in Equation~(\ref{eq:likelihoodNaive}) %for some reason this says 1.22 in my pdf build and not 1.21 as it should be, or is this just due to my TeX environment?
to the probability of observing the data $TCACC$ for the tips and $x_{0},x_{6},x_{7},x_{8}$ for the ancestral nodes (inside the brackets) and the sum over all possible path combinations over the state space $\mathcal{E}$ (outside the brackets).
Term $\pi_{0}$ is simply the probability that the root has character $x_{0}$ and is based on the assumption that the process generating the data has reached its equilibrium.

Summing over all possible paths leading to the observed data is computationally expensive, as there are  $K^{n-1}$ possible combinations, where $n-1$ is the number of internal nodes of a rooted binary tree.
Fortunately, we can rely on the assumption of independence of the process in the two sub-trees below every node.
Based on this, \citet{Felsenstein1981} provides an efficient algorithm for computing the likelihood of observing the data on a topology $\mathbf{F}$ given the substitution model that quantifies transition probabilities.
This routine, known as \emph{tree pruning} conveniently expresses the partial likelihood $L_{i}(x_{i})$ of observing data at the descendants of node $i$ given state $x_{i}$ at node $i$ in terms of partial likelihoods at nodes $j$ and $k$.
For all internal nodes of $\mathbf{F}$ we have:

\begin{equation}
L_{i}(x_{i})=\left[\underset{x_{j}\in \mathcal{E}}{\sum}\mathbf{P}_{x_{i}x_{j}}(t_{j})L_{j}(x_{j})\right]\times\left[\underset{x_{k}\in \mathcal{E}}{\sum}\mathbf{P}_{x_{i}x_{k}}(t_{k})L_{k}(x_{k})\right]
\label{eq:felsenstein1}
\end{equation}

\noindent
For tip nodes we have:

\begin{equation}
L_{i}(x_{i})=\begin{cases}
1 & \text{if }\hat{x}(i)=x_{i}\\
0 & \text{otherwise}
\end{cases},
\label{eq:felsenstein2}
\end{equation}

\noindent
where $\hat{x}(i)$ denotes the character state observed at node $i$.
After recursively applying (\ref{eq:felsenstein1}) to all the nodes of $\mathbf{F}$ in post-order fashion the likelihood of the data at the site is given by:
% we can integrate out the unobserved states calculating successive contributions to the partial likelihood.

\begin{equation}
L(\mathbf{x}_{h})=\underset{x_{0}\in\mathcal{E}}{\sum}\pi_{x_{0}}\cdot L_{0}(x_{0})
\label{eq:felsenstein3}
\end{equation}

\noindent
Thanks to the assumption of site-independence, the log-likelihood of the whole alignment $X$ can be computed by summing the logarithms of the likelihoods at the particular sites:

\begin{equation}
l(X)=\underset{j=1}{\overset{l}{\sum}}log\left(L(\mathbf{x}_{jh})\right)
\label{eq:loglikelihood}
\end{equation}

%PL: perhaps note that you moved to Log L in the last step to avoid underflow when multiplying small numbers or something?

\subsection{The computational burden of likelihood computations\label{sub:mlburden}}

We have already mentioned that for large data sets the run-time of statistical phylogenetic analysis is a major caveat that hampers the development and application of novel, more biologically relevant models and inference in the field.
These limits are being actively stretched, for example by porting the computationally expensive calculations to high-performance parallel devices such as Graphics-Processing Units (GPUs, \cite{Nickolls2008}).
In Subsection~\ref{sub:exponentiation} we already discussed how the computation of finite-time transition probabilities can be reduced to simple matrix algebra operations, which can handled by high-performance libraries and APIs for statistical phylogenetics such as the BEAGLE library \citep{Ayres2012, Suchard2009}.

%GB: do we need this computational complexity section? Not sure if we want to go into that much detail ...
Before we proceed with discussing the burden imposed by certain steps when calculating the likelihood of a sequence evolution on a tree we need a formal definition of the computational complexity of a sequential routine.
The efficiency of an algorithm can depend on various factors, like network usage, disk usage, operational memory usage or CPU real-time usage.
All of these factors should be taken into consideration when designing and implementing a computational routine. 
However, for most of the calculations the CPU time will be the rate-limiting factor.
Typically both mathematicians and computer scientists use the Big-O notation as a relative representation of algorithm complexity, like the execution time required.
Here we formally define the Big-O notation as:

\begin{definition}
Let us assume two functions $f:\mathbb{R}_{>0}\rightarrow\mathbb{R}_{>0}$ and $T:\mathbb{R}_{>0}\rightarrow\mathbb{R}_{>0}$. 
For $n \in \mathbb{R}_{>0}$ we write:

\begin{equation}
T(n) \in \mathcal{O}\left(f(n)\right)
\label{eq:bigOh}
\end{equation}

\noindent
if and only if there exists a constant $M>0$ and some $n_0 \in \mathbb{R}_{>0}$ such that:

\begin{equation}
T(n)\leq M \cdot f(n)
\end{equation}
\end{definition}

% \noindent
In other words $T(n)$ is $\mathcal{O}\left(f(n)\right)$ if and only if for all sufficiently large input values $T(n)$ is bounded from above by some constant multiple of $f(n)$.
This makes the Big-O notation useful when analyzing algorithms for run-time efficiency and makes it possible to, for example, compare routines without the added overhead of negligible terms, such as the architecture at which the code was run, the choice of language or compiler.
We will now briefly discuss some of the most common serial computational orders and their examples.

\begin{algorithm}[H]
\centering
\begin{algorithmic}[1]
% \footnotesize{
\State $a \gets \text{\textbf{new int}}$;
%
\If{ $\left(a == 1 \right)$ }
%
\State \textbf{return} $true$;
%
\Else 
%
\State \textbf{return} $false$;
%
\EndIf
% }
\end{algorithmic}
\caption{
{ \footnotesize 
{\bf Single test operation.} 
} % END: footnotesize
}
\label{alg:o1}
\end{algorithm}

The simplest, most basic example is an operation that takes the same amount of time every time it is called, thus requiring constant time, i.e. it's computational complexity is that of $\mathcal{O}\left(1\right)$ and does not depend on the input size.
An example of such routine is listed in the Algorithm~\ref{alg:o1} that evaluates a specific value of a constant.

\begin{algorithm}[H]
\centering
\begin{algorithmic}[1]
% \footnotesize{
\State $a \gets \text{\textbf{new int}}\left[n\right]$;
%
\For{int $\left(i=0; \; i<a.size(); \; i++\right)$}
%
\If{ $\left(a\left[i\right] == 1 \right)$ }
%
\State \textbf{return} $true$;
%
\Else 
%
\State \textbf{return} $false$;
%
\EndIf
%
\EndFor
% }
\end{algorithmic}
\caption{
{ \footnotesize 
{\bf Search for a value.} 
}% END: footnotesize
}
\label{alg:on}
\end{algorithm}

An algorithm proceeding at $\mathcal{O}\left(N\right)$ has a run-time that grows linearly with the size of input data.
An example of such a routine is presented as Algorithm~\ref{alg:on}, which searches an array looking for a specific value.
We should note that the Big-O notation always describes the performance in worst-case scenario, i.e. a full loop has to be executed before a matching element is found, while for the actual real-world problem this does not have to be the case and the loop can return earlier.

\begin{algorithm}[H]
\centering
\begin{algorithmic}[1]
% \footnotesize{
\State $a \gets \text{\textbf{new int}}\left[n\right]$;
%
\For{int $\left(i=0; \; i<a.size(); \; i++\right)$}
%
\For{int $\left(j=0; \; j<a.size(); \; j++\right)$}
%
\If{ $\left(i == j \right)$ }
%
\State \textbf{do nothing};
%
\Else 
%
\If{ $a\left[ i \right] == a\left[ j \right]$ }
%
\State \textbf{return} $true$;
%
\EndIf
%
\EndIf \Comment{END: i == j check}
%
\EndFor \Comment{END: j loop}
%
\EndFor \Comment{END: i loop}
% }
\end{algorithmic}
\caption{
{ \footnotesize 
{\bf Search for a first duplicated value.} 
}% END: footnotesize
}
\label{alg:on2}
\end{algorithm}

An algorithm can also require a quadratic run-time, $\mathcal{O}\left(N^2\right)$, i.e. its performance is directly proportional to the square of the size of the input data.
Among examples of such an algorithm is the routine that searches for a duplicated value in an array, and requires two nested loops as presented in Algorithm~\ref{alg:on2}.
Similarly nesting more loops will result in routines which proceed at polynomial times $\mathcal{O}\left(N^3\right)$, $\mathcal{O}\left(N^4\right)$, $\mathcal{O}\left(N^5\right)$, etc.
An algorithm proceeding at $\mathcal{O}\left(c^N\right)$, where $c>1$ is some constant, will have a run-time that grows c-times with every additional data element, i.e. exponentially.

Having described these basic concepts, we can now come back to the original problem of calculating the likelihood of sequence evolution on a fixed tree.
From Subsections \ref{sub:exponentiation} and \ref{sub:likelihood} we already know the steps involved:

% I skipped site rate variability, but perhaps it should be mentioned? 
\begin{enumerate}
\item { Singular value decomposition of the rate matrix $\mathbf{Q}$. }
\item { Taking the exponential of $\mathbf{Q}t$ for each branch length $t$. }
\item { Applying Equation~(\ref{eq:felsenstein1}) for every site in the alignment and every possible state in the state space $\mathcal{E}$. }
\item { Taking the logarithm and summing over all sites in the alignment. }
\end{enumerate}

Step (i) proceeds at ${\cal{O}}(K^3)$; however, for typical, time-homogeneous applications this is not the a rate-limiting step, as repeated evaluations of $\exp(\mathbf{Q}t)$ in (ii) are based on the same decomposition of the matrix \citep{Suchard2009}. 
Step (ii) typically boils down to repeated matrix-vector-matrix multiplications and its serial computational order is ${\cal{O}}(K^3 \times n)$. 
Step (iii) takes ${\cal{O}}(K^2 \times n \times l)$ time and is the most computationally expensive part of the routine. 
Step (iv) proceeds at ${\cal{O}}(l)$, linearly with the number of sites.
Fortunately, these operations are very regular and lend themselves to the computational parallelism, such as utilized in the BEAGLE library \citep{Suchard2009,Ayres2012}.

% TODO: maybe sub-section on GPUs here

\section{Bayesian inference\label{sec:bayesian_inference}}

\subsection{Bayesian evolutionary analysis}

The Bayesian approach to statistical inference combines the prior information with the data to generate the posterior distributions of the parameters of interest.
All other post-hoc inference is based on those generated posterior distributions.
Bayesian methodology has recently gained immense popularity, due to the advances in numerical algorithms and the progressively cheaper access to powerful computational hardware.
The development of statistical programs for simulation from Bayesian hierarchical models like OpenBUGS \citep{Lunn2009} and JAGS \citep{Plummer2003} has also helped to bolster the popularity and accessibility of the Bayesian framework.

In the field of phylogenetics, Bayesian inference has been applied to some of the most complex problems and is now widely adopted, with the term \emph{Bayesian evolutionary analysis} used to describe the molecular evolutionary analysis set in the Bayesian framework.
At the fore-front of these advancements lies the active development and wide adoption of popular Bayesian evolutionary analysis software packages like MrBayes \citep{Ronquist2012} and BEAST \citep{Drummond2012}.

\subsection{Bayes theorem\label{sub:bayesTheorem}}

% Following the notation defined in Subsection \ref{sub:likelihood} we will denote the tree topology, the branch lengths and the parameters of the substitution model that we aim at inferring by $\mathbf{\Theta}$.
Following the notation defined in Subsection~\ref{sub:likelihood}, let us denote a single-dimensional parameter of interest that we aim to learn about as $\theta$.
The central concept of Bayesian inference is that this parameter has a distribution of its own.
Before the molecular data $\mathbf{X}$ is observed we assume that $\theta$ has a prior distribution $P(\theta)$.
The prior reflects the prior beliefs about the possible values of the parameter.

The prior choice is perhaps one of the most controversial points of the Bayesian inference.
The prior distributions may come from the preceding analysis of the other sets of data, or they may form some set of biologically plausible values.
However in some cases there may be no suitable objective distribution at hand, which case one often resorts to the so called \emph{vague} priors, with large variance.
All of these proper prior choices lead to a proper posterior distribution, yet some controversies about their choice may still arise.

% TODO: discuss hyperpriors?
The prior information is then combined with the likelihood of the data $P\left(\theta|\mathbf{X}\right)$ to form the posterior, via the Bayes theorem (Equation~(\ref{eq:bayesian1})).
The posterior distribution of $\theta$ follows:

\begin{eqnarray}
P\left(\theta|\mathbf{X}\right) &=& \frac{P(\theta)\cdot P\left(\mathbf{X}|\theta\right)}{P\left(\mathbf{X}\right)}=\frac{P(\theta)\cdot P\left(\mathbf{X}|\theta\right)}{\int P(\theta)\cdot P\left(\mathbf{X}|\theta\right)d\theta} \\ \nonumber
& \approx & P(\theta)\cdot P\left(\mathbf{X}|\theta\right),
\label{eq:bayesian1}
\end{eqnarray}

\noindent
where the marginal probability of the data $P\left(\mathbf{X}\right)$ is a normalizing constant that makes the posterior $P\left(\theta|\mathbf{X}\right)$ integrate to one.
The last approximation in Equation~(\ref{eq:bayesian1}) means that the posterior is proportional to the prior times the likelihood.

% nuisance parameters? GB: I second that
Bayesian inference provides a natural way of dealing with nuisance parameter.
Suppose that $\mathbf{\Theta}=\left(\theta,\mathbf{F}\right)$ is now a two dimensional vector of parameters, with $\theta$ being the parameter of interest, while 
$\mathbf{F}$ (e.g. a tree topology) is the nuisance parameter.
The joint posterior distribution of $\mathbf{\Theta}$ is now:

\begin{eqnarray}
P\left(\mathbf{\Theta}|\mathbf{X}\right) &=& P\left(\theta,\mathbf{F}|\mathbf{X}\right)=\frac{P\left(\theta,\mathbf{F}\right)\cdot P\left(\mathbf{X}|\theta,\mathbf{F}\right)}{P\left(\mathbf{X}\right)} \\ \nonumber
&=& \frac{P\left(\theta,\mathbf{F}\right)\cdot P\left(\mathbf{X}|\theta,\mathbf{F}\right)}{\int\int P(\theta,\mathbf{F})\cdot P\left(\mathbf{X}|\theta,\mathbf{F}\right)d\theta d\mathbf{F}}
\label{eq:bayesian2}
\end{eqnarray}

\noindent
and the marginal posterior density of $\theta$ can be obtained from Equation~\ref{eq:bayesian2} as:

\begin{equation}
P\left(\theta|\mathbf{X}\right) =\int P\left(\mathbf{\Theta},\mathbf{F}|\mathbf{X}\right)d\mathbf{F}
\label{eq:bayesian3}
\end{equation}

\subsection{Example: Comparing two sequences with the beta-binomial model}

As a simple illustration of the concept of Bayesian analysis, let us consider a toy example where two sequences are compared and $x=10$ differences are found out of $N=100$ sites compared.
Suppose that the parameter $\theta$ we want to infer is the proportion of sites which differ between the two sequences.
Assuming that the sites are independent we can use a binomial model and put a beta prior with parameters $\alpha=2$, $\beta=100$ on the unknown proportion $\theta$. %PL: maybe add explicitly: ".., which prefers small values of the $\theta$ a priori

%PL: to avoid color, use a white filled distribution for the prior, a black one for the likelihood, and a grey for the posterior (which conceptually fits with the fact that the posterior combines black and white)

%---LIKELIHOOD---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{bayes} 
\caption{
{ \footnotesize 
{\bf Beta prior and posterior densities plotted with a scaled likelihood resulting from a binomial model.} 
}% END: footnotesize
}
\label{fig:bayes1}
\end{figure}

Figure~\ref{fig:bayes1} shows the resulting posterior distribution along with prior and scaled likelihood.
This illustrates that the data `pulls' the posterior away from the prior, and the strength by which is does so depends on its variance, resulting in a posterior centered around 0.6 in this case.
In this example, the posterior distribution can be obtained analytically because the beta distribution is a so called \emph{conjugate prior} for the binomial model, resulting in the posterior that is also beta-distributed.

\subsection{Markov Chain Monte Carlo\label{sub:mcmc} }

Except for the most trivial problems, involving conjugate priors as in our previous example, the calculation of the integral in Equation~(\ref{eq:bayesian1}) poses a considerable challenge in Bayesian inference.
This integral leads to the marginal probability of the data $P\left(\mathbf{X}\right)$. 
For even the simplest phylogenetic problems, one has to integrate over all possible tree topologies, all of the branch lengths in those trees and all of the substitution parameters in the applied model.
Such highly-dimensional integrals are impossible to calculate analytically.
This is where the Markov Chain Monte Carlo (MCMC) algorithm shows its strength and provides a powerful method of sampling from the posterior distributions without evaluating the marginal probability. %PL: it is a supa-dupa algorithm: http://oldweb.cecm.sfu.ca/~jborwein/algorithms.html

The Metropolis-Hasting algorithm, as proposed by \citet{Metropolis1953}, is an MCMC algorithm for sampling from multi-dimensional distributions.
It proceeds by constructing a Markov chain with its states being the values of the parameter of interest $\theta$ and whose equilibrium distribution is the posterior $P\left(\theta|\mathbf{X}\right)$ as defined in Equation~\ref{eq:bayesian1}.
After choosing an arbitrary starting state for the chain, new candidate states $\theta^{*}$ are generated from the proposal density $q(\theta^{*} | \theta[t])$ also called the kernel.

A common kernel choice to propose new state for continuous parameter is to sample from a uniform distribution centered around the current state of the chain, the so called jumping kernel: %PL: are you sure this sliding window kernel is commonly referred to as the 'jumping kernel", or do you mean as a general term?

$$\theta^{*}\sim U\left[\theta[t]-\omega,\theta[t]+\omega\right],$$

\noindent
where $\omega$ is the arbitrary window size.
The chain then moves to the proposed value with probability defined by the acceptance ratio (see Equation~(\ref{eq:metropolis1})), or if the proposal is rejected the chain remains at the current state.
The routine iterates between accept/reject steps for a specified number of times.
Both acceptance and rejection are counted as iterations of the algorithm; in fact the rejections ensure that the probabilistic properties of the target distribution are well represented in the generated sample.  
Algorithm~\ref{alg:metropolisHastings} presents the details of the Metropolis-Hastings algorithm.

\begin{algorithm}[H]
\centering
\begin{algorithmic}[1]
% \footnotesize{
%
\State Pick a starting value of the chain $\theta \left[ t \right]$ for $t=0$.
%
\For{int $\left(t=0; \; t<\text{chain length}; \; t++\right)$}
%
\State Generate a candidate state from the proposal distribution (kernel):
$$\theta^{*} \sim q(\theta[t])$$
%
\State Calculate the acceptance ratio:
$$r=\frac{prior\left(\theta^{*}\right)}{prior\left(\theta\left[t\right]\right)}\cdot\frac{Likelihood\left(\mathbf{X}|\theta^{*}\right)}{Likelihood\left(\mathbf{X}|\theta\left[t\right]\right)}\cdot\frac{q\left(\theta\left[t\right]|\theta^{*}\right)}{q\left(\theta^{*}|\theta\left[t\right]\right)}.$$
%
\State Sample $u\sim U[0,1]$.
%
\If{$\left( u < r \right)$}
%
\State Set $\theta[t+1]=\theta^*.$
%
\Else 
%
\State Set $\theta[t+1]=\theta[t].$
%
\EndIf
%
\EndFor
% }
\end{algorithmic}
\caption{
{ \footnotesize 
{\bf The Metropolis-Hastings algorithm} 
}% END: footnotesize
}
\label{alg:metropolisHastings}
\end{algorithm}

Several important features of the algorithm can be noted.
The most important part of the algorithm is the proposal density. 
Given the current state it proposes the next state independent of the past states.
We can recall that this is the Markov property, in fact the output of the algorithm generates a Markov chain, hence the name MCMC.
The Metropolis-Hastings algorithm allows for the use of an asymmetrical proposal densities such that $q\left(\theta\left[t\right]|\theta^{*}\right)\neq q\left(\theta^{*}|\theta\left[t\right]\right)$.
The acceptance probability is simply the ratio of the target posterior evaluated at the proposal and at the current state of the chain, times the ratio of the proposal: 
% prior ratio times the likelihood ratio times the proposal ratio

\begin{equation}
\frac{P\left(\theta^{*}|\mathbf{X}\right)\cdot q\left(\theta\left[t\right]|\theta^{*}\right)}{P\left(\theta\left[t\right]|\mathbf{X}\right)\cdot q\left(\theta\left[t\right]|\theta^{*}\right)}=\frac{P(\theta^{*})\cdot P\left(\mathbf{X}|\theta^{*}\right)\cdot q\left(\theta\left[t\right]|\theta^{*}\right)}{P(\theta\left[t\right])\cdot P\left(\mathbf{X}|\theta\left[t\right]\right)\cdot q\left(\theta^{*}|\theta\left[t\right]\right)}.
\label{eq:metropolis1}
\end{equation}

Note that the integral in Equation~(\ref{eq:bayesian1}) which is notoriously hard to compute cancels out in (\ref{eq:metropolis1}).
Thus to estimate the posterior $P\left(\theta|\mathbf{X}\right)$ one only needs to run the algorithm for a sufficiently long time to let the chain reach it's equilibrium, which is precisely the target posterior.
%PL: maybe mention that this can be formally proven using the ergodic theorem... and no, you don't have to prove it in the appendix ;-)

% TODO: better title, show how to derive the probabilities (in JC69 chapter)
\subsection{Example: Estimating genetic distance}

%GB: I don't really like these kinds of examples or tutorials. They're great when teaching a workshop, but this is an introduction for a PhD dissertation ...

Let us revisit the toy example in which we compare two sequences and find $x=10$ differences among $n=100$ sites.
This time however we will use MCMC to estimate the genetic distance between the two sequences using the Jukes and Cantor model of nucleotide substitution (see Subsection~\ref{sub:jc69}).
Genetic distance is defined as the expected number of substitutions per nucleotide site, which we will denote as $\theta=rt$ in line with previous sections, where $r$ is the substitution rate (see Subsection~\ref{sub:poisson}) and $t$ is the evolutionary time separating the two sequences.
The time and rate are generally confounded as we already mentioned, and hence they will be estimated as a single parameter of interest $\theta$.
The probability that the descendant nucleotide in a sequence is different from the ancestral nucleotide can be obtained by summing over all possible paths, between two pairs of nucleotides, i.e. $p_{ij}(t)=\underset{k\neq i}{\sum}p_{ik}(t)$.

From Equation~(\ref{eq:jc69Finite}) in Subsection~\ref{sub:jc69}, we know that under the JC69 model those probabilities are all equal, therefore the probability that a single site is different between two sequences separated by distance $\theta=rt$ is given by:

\begin{equation}
p=3p_{1}\left(t\right)=\frac{3}{4}\left(1-e^{\left(-4/3\right)\theta}\right).
\label{eq:distance1}
\end{equation}

\noindent
From this we can formalize the likelihood function using the binomial model given that we observe $x$ differences out of $n$ sites :

\begin{equation}
P\left(\mathbf{X}|\theta\right)=L\left(\mathbf{X}|\theta\right)=\left(\begin{array}{c}
n\\
k
\end{array}\right)p^{x}(1-p)^{n-x},
\label{eq:likelihood1}
\end{equation}

\noindent
where $p$ is as defined by Equation~(\ref{eq:distance1}).
We use an exponential prior on $\theta$ with mean $1/\lambda$:

\begin{equation}
P\left(\theta,\lambda\right)=\frac{1}{\lambda}e^{-(1/\lambda)\theta},
\label{eq:expPrior}
\end{equation}

\noindent
and set $\lambda=10$.
For the proposal we choose a uniformly distributed jumping kernel with a window size of $w=0.1$.

%---METROPOLIS PATH---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{metropolis} 
\caption{
{ \footnotesize 
{\bf MCMC output for estimating genetic distance under the JC69 model of nucleotide substitution.
%PL: a bit thin as a caption; at least describe what we see in the two panels (in a slightly different way as the text) -- Figures should be able to be read in isolation
%PL: use a grey line instead of the two red lines to avoid color.
} 
}% END: footnotesize
}
\label{fig:metropolis}
\end{figure}

Figure~\ref{fig:metropolis} shows $1000$ iterations of the chain generated by Algorithm~\ref{alg:metropolisHastings} in the upper row and applied to the genetic distance problem.
The red line 
%PL: let's make it grey
indicates the maximum likelihood estimate (MLE) $\hat{\theta}$ of parameter $\theta$ which is calculated by setting $\frac{d\ell\left(\mathbf{X}|\theta\right)}{d\theta}=0$, where $l\left(\mathbf{X}|\theta\right)$ is the logarithm of likelihood defined by Equation~(\ref{eq:likelihood1}):

\begin{equation}
\hat{\theta}=\frac{3}{4}log\left(1-\frac{4}{3}\cdot\frac{x}{n}\right)
\label{eq:mle}
\end{equation}

In the second row of Figure~\ref{fig:metropolis} we can see an estimate of the posterior density, along with the sample mean.
The posterior sample mean is identical to the MLE up to the second decimal place, although we have not run the chain for long.
The acceptance rate is high, suggesting that almost as many steps are accepted as are rejected.
%GB: acceptance rates don't necessarily have to be this high, not sure if you want to discuss this though.

%PL: you could potentially mention the importance of tuning kernels, like the window size, and show the same trace plot for a  too high and too low window, and their acceptance ratio. This can lead to a mention of diagnozing mcmc runs, convergence, ESSs, and shit... which is a practical issue for complex problems/large data sets, etc.
%GB: I wouldn't discuss tuning, but acceptance ratios I think are more important.

\subsection{Bayesian phylogeography\label{sub:phylogeo}}

Phylogeography is a research direction aimed at connecting evolutionary processes, which happen over time, with the processes of spatial distribution, in actual geographical coordinates, into a joint spatio-temporal dynamic that can be used to track pathogen spread.
Historically the phylogeographical inference in molecular epidemiology ignored the reciprocal character of these two processes. 
The evolutionary relationship represented by the tree was inferred omitting the geographical information, and then the ancestral locations generated by the process of the spatial diffusion were reconstructed, conditioning on the inferred tree topology.
\citet{Lemey2009} first considered the Bayesian reconstruction that allows for joint inference of both the tree and the parameters of the spatial and temporal process that shapes the spread and evolution of 
% e.g. 
rapidly evolving viruses.
Their approach does not need to fix the tree topology, the branch lengths nor the Markov model parameters, and allows for integration over all possible values, requiring only for a specification of a prior distributions for the parameters that we aim to infer.

%TODO: plot, tree with locations
%---BAYESIAN PHYLOGENETICS TREE---%
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{phylogeotree} 
\caption{
{ \footnotesize 
{\bf Spatial location and molecular sequence data observed at the tips of a phylogenetic tree.} 
}% END: footnotesize
}
\label{fig:phylogeotree}
\end{figure}

Central to the Bayesian phylogeographic framework is the hypothesis that sequence evolution is occurring simultaneously with the process of geographical dispersal. 
The observed data, residing at the $N$ tips of phylogeny $\mathbf{F}$, is recorded as both character sequence data $\mathbf{X}=(X_{1},...,X_{N})$ and discretized geographic locations (countries, districts, cities) $\mathbf{Y}=(Y_{1},...,Y_{N})$.
We assume independent stochastic processes are responsible for generating these data, with ancestral states of $\mathbf{X}$ generated as before by a CTMC characterized by rate matrix $\mathbf{Q}$ and the unobserved ancestral locations $(Y_{N+1},...,Y_{2N-2})$ generated by a separate CTMC characterized by rate matrix $\mathbf{\Phi}$.
The entries of this matrix are the transition rates $\phi_{ij}$ between locations $i$ and $j$.
Unlike before the rate matrix does not necessarily have to be symmetrical, i.e. there exist some $i,j$ such that $\phi_{ij}\neq\phi_{ji}$.
Additionaly, a priori one would expect many of the infinitesimal rates to be zero. 
Unlike the molecular data, which is typically dense, the location data is sparse, with only one observation per taxon and many connections not observed at all.

The Bayesian framework described in Subsections~\ref{sub:bayesTheorem} and \ref{sub:mcmc} offers the unique ability to integrate different sources of information about the viruses with their genetic data, without conditioning on either one of them.
We use this ability to infer the tree topology, the ancestral spatial locations and the character substitution process giving rise to the aligned molecular sequence data. 
We assume independent stochastic processes are responsible for generating these data.
This allows us to write down the joint posterior distribution:

% \substack{\text{aaa} \\ \text{bbb}}

\begin{eqnarray}
% \footnotesize{
P\left(\mathbf{F}, \mathbf{Q}, \mathbf{\Phi}|\mathbf{X},\mathbf{Y}\right)\propto 
\underbrace{P(\mathbf{X}|\mathbf{Q}, \mathbf{F})}_{\begin{array}{c}
\substack{\text{Sequence} \\ \text{likelihood}}
\end{array}}\cdot\underbrace{P(\mathbf{Y}|\mathbf{\Phi}, \mathbf{F})}_{\begin{array}{c}
\substack{\text{Trait} \\ \text{likelihood}}
\end{array}}\cdot
\underbrace{P(\mathbf{F})}_{\begin{array}{c}
\substack{\text{Topology} \\ \text{prior}}
\end{array}} \\ \nonumber
\cdot\underbrace{P(\mathbf{Q})}_{\begin{array}{c}
\substack{\text{Sequence substitution} \\ \text{prior}}
\end{array}}\cdot\underbrace{P(\mathbf{\Phi})}_{\begin{array}{c}
\substack{\text{Location exchange} \\ \text{prior}}
\end{array}}
% }
\label{eq:posteriorPhylogeography}
\end{eqnarray}

The likelihood component for both molecular data $\mathbf{X}$ and location data $\mathbf{Y}$ is calculated using recursive tree pruning, as described in Subsection~\ref{sub:likelihood}.
We approximate the joint posterior as written down in Equation~(\ref{eq:posteriorPhylogeography}) using MCMC methods we talked about in Subsection~\ref{sub:mcmc}.