### 2014-11-14 frontback/acknowledgments.tex
summary

% Acknowledgements ============================================================

\pdfbookmark[1]{Acknowledgments}{Acknowledgments}
\chapter*{Acknowledgments}

TODO

\vfill{}

The research leading to these results has received funding from the European Research Council under the European Community's Seventh Framework Programme (FP7/2007-2013) under Grant Agreement no. 278433-PREDEMICS and ERC Grant agreement no. 260864, the Welcome Trust (grant no. 092807) and the National Institutes of Health (R01 GM086887 and R01 HG006139) and the National Science Foundation (IIS 1251151 and DMS 1264153).
The National Evolutionary Synthesis Center (NESCent) catalyzed this collaboration through a working group (NSF EF-0423641). LMC would like to thank the Program for Scientific Computing staff for operational support.