### 2014-11-14 frontback/disclaimer.tex

% Disclaimer ====================================================================

\vspace*{\fill}
\begin{center}
{\it This manuscript has been submitted to the Department of Biomedical Sciences 
in partial fulfillment of the requirements for the degree of Doctor of Biomedical Sciences.

\vskip1cm

% TODO: comment out
This version of the manuscript is pending the 
%approval of the big boss Philippe.
review by the jury members
}%END: it
\end{center}
\vspace*{\fill}
\vspace*{\fill}
\vspace*{\fill}