### 2020-12-11 frontback/jury.tex
last minute modifications

% Jury ====================================================================

\pdfbookmark[1]{Jury members}{Jury members}
\chapter*{Jury members}

\noindent \textsc{David Posada}, Professor of Genetics at the University of Vigo, Vigo, Spain; \\

\noindent \textsc{Stijn Vansteelandt}, Professor of Statistics at the Ghent University, Ghent, Belgium \\

\noindent \textsc{Stein Aerts}, Professor of Bioinformatics at the University of Leuven, Leuven, Belgium \\

\noindent \textsc{Geert Molenberghs}, Professor of Biostatistics at the University of Leuven, Leuven, Belgium \\

\noindent \textsc{Peter Hoet}, Professor of Pneumology at the University of Leuven, Leuven, Belgium \\